\documentclass{report}
\usepackage{amsmath,amssymb}

\def\mathbi#1{\textbf{\em #1}}
\def\BA{\bf{A}}
\def\BB{\bf{B}}
\def\bn{\bf{n}}
\def\CP{\mathcal{P}}
\def\CV{\mathcal{V}}
\def\CI{\mathcal{I}}

\newcommand{\gradt}{\nabla\cdot}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\Cplex}{\mathbb{C}}

\newcommand{\Vecc}[2]{ \left(
	\begin{array}{c}
	#1 \\ #2
	\end{array}
	\right) }

\newcommand{\Matrr}[4]{ \left(
	\begin{array}{cc}
	#1 & #2 \\
	#3 & #4
	\end{array}
	\right) }

\newcommand{\calB}{{\mathcal B}}
\newcommand{\calBB}{{\mathcal B}^{\Box}}
\newcommand{\calK}{\mathcal{K}}
\newcommand{\bfA}{{\mathbf A}}
\newcommand{\bfb}{{\mathbf b}}
\newcommand{\bfr}{{\mathbf r}}
\newcommand{\bfx}{{\mathbf x}}
\newcommand{\bfxex}{{\mathbf x}_{\mbox{\scriptsize $\star$}}}
\newcommand{\Dim}{\mathop{\mathrm{dim\ }}}

\begin{document}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{center}
{\large
{\bf Improving coarsening and interpolation for algebraic multigrid}}

	Jeff Butler \\
	Dept.~of Applied Mathematics, University of Waterloo \\
	Waterloo, Ontario, Canada N2L 3G1 \\
	{\tt jsbutler@math.uwaterloo.ca} \\
	Hans De Sterck
\end{center}
Algebraic multigrid (AMG) is one of the most efficient
algorithms for solving large sparse linear systems on
unstructured grids. Classical coarsening schemes such as the
standard Ruge-Steuben method [1] can lead to adverse effects
on computation time and memory usage that affect
scalability. Memory and execution time complexity growth is
remedied for various large three-dimensional (3D) problems
using the parallel modified independent set (PMIS)
coarsening strategy developed by De Sterck, Yang, and Heys
[2]. However, this coarsening strategy often leads to
erratic grids without a regular structure that diminish
convergence. This talk looks at a modification of the PMIS
algorithm that consistently produces more uniform
coarsenings, and significantly improves convergence
properties over the original PMIS algorithm. Improvements of
existing interpolation schemes and application of the
resulting methods to several problems are also examined.

[1] J.~Ruge, K.~Stueben, {\em Algebraic multigrid (AMG)},
Multigrid Methods v.3, Frontiers in Applied Mathematics,
S.~McCormick, ed., SIAM (1987) 73--130.

[2] H.~De Sterck, U.~M.~Yang, and J.~J.~Heys,
{\em Reducing Complexity in Parallel Algebraic
Multigrid Preconditioners}, to appear in
SIAM J.~Matr. An. and Applications, 2006.



	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\end{document}

