\documentclass{report}
\usepackage{amsmath,amssymb}
\setlength{\parindent}{0mm}
\setlength{\parskip}{1em}
\begin{document}
\begin{center}
\rule{6in}{1pt} \
{\large Dirk Abbeloos \\
{\bf Nested multigrid methods for time-periodic, parabolic optimal control problems.}}

K U Leuven \\ Celestijnenlaan 200A \\ 3001 Heverlee
\\
{\tt dirk.abbeloos@cs.kuleuven.be}\\
Stefan Vandewalle \\
Michael Hinze\\
Moritz Diehl\end{center}

We present a nested multigrid method to optimize time-periodic,
parabolic, partial differential equations (PDE). We consider a quadratic
tracking objective with a linear parabolic PDE constraint. The goal of
the optimization is to find the optimal control which minimizes the
objective over a set of admissible controls and in addition induces a
prescribed cyclic state with the correct period. More precisely the state
of the system at a certain instance of time should be identical to the
one a fixed period of time later. We do not assume the existence of
autonomous periodic behavior of the PDE and as a consequence the
time-periodicity should be enforced by the applied control. Our problem
differs from the initial value problem, where the goal is to find the
optimal control to steer the system from a given state to a final state.

The proposed method relies on the reduction of the first order optimality
conditions to a Fredholm integral equation of the second kind. The
resulting equation will be solved by a multigrid method of the second
kind. The cost of this method is dominated by the evaluation of the
kernel of the integral, which involves solving the state and the
corresponding adjoint equation. For the evaluation of these two
subproblems a space-time multigrid approach of the first kind will be
applied. Combining both multigrid methods results in a nested multigrid
algorithm where an outer method is used for solving the integral equation
and an inner method to evaluate the integral kernel.


\end{document}
