\documentclass{report}
\usepackage{amsmath,amssymb}
\setlength{\parindent}{0mm}
\setlength{\parskip}{1em}
\begin{document}
\begin{center}
\rule{6in}{1pt} \
{\large Justin, W.L. Wan \\
{\bf Multigrid Methods for Solving Hamilton-Jacobi-Bellman and Isaacs Equations Arising from Nonlinear Option Pricing Problems}}

200 University Avenue West \\ University of Waterloo \\ Cheriton School of Computer Science \\ Waterloo \\ Ontario N2L 3G1 \\ Canada
\\
{\tt jwlwan@uwaterloo.ca}\\
Dong Han\end{center}

Many financial pricing problems can be formulated as optimal control
problems, leading to nonlinear Hamilton-Jacobi-Bellman (HJB) partial
differential equations (PDEs) or Hamilton-Jacobi-Bellman-Isaacs (HJBI)
equations. In general, the HJB equations can be written as:
$$
%-mb: \begin{eqnarray*}
V_\tau = \sup_{q \in Q} \left\{
a(S,\tau,q) V_{SS} + b(S,\tau,q) V_S - c(S,\tau,q) V \right\},
%-mb: \end{eqnarray*}
$$
where $V$ typically represents the value of an option, $S$ is the
underlying asset price, $\tau$ is the time to expiry, and $q$ is a
control parameter. In the case of HJBI equations, they can be written as:
$$
%-mb: \begin{eqnarray*}
V_\tau = \sup_{q \in Q} \inf_{p \in P} \left\{
a(S,\tau,q,p) V_{SS} + b(S,\tau,q,p) V_S - c(S,\tau,q,p) V \right\},
%-mb: \end{eqnarray*}
$$
where $p$ is an additional control parameter. These two types of equations
are highly nonlinear.

An implicit discretization yields a nonlinear set of equations which must
be solved at each time step. Standard methods such as fixed point
iterations can be very slow for fine grids. In the case of HJB equations,
a Newton-type iteration scheme, often known as the Policy iteration in
stochastic control, has been proposed. Quadratic convergence is observed
when close to solution. However, in the case of HJBI equations, Policy
iteration may not necessarily converge.

In this talk, we present multigrid methods for solving the HJB and HJBI
equations. The use of Policy iteration and FAS multigrid will be
discussed. The proposed multigrid methods are based on a relaxation
scheme as smoother which is convergent for both HJB and HJBI equations.
We show by Fourier analysis the smoothing property of the relaxation
scheme. Linear interpolation and full weighting are used for intergrid
transfer. We demonstrate numerically the effectiveness of the multigrid
methods by examples of financial pricing problems.


\end{document}
