\documentclass{report}
\usepackage{amsmath,amssymb}
\setlength{\parindent}{0mm}
\setlength{\parskip}{1em}
\begin{document}
\begin{center}
\rule{6in}{1pt} \
{\large Luis Chacon \\
{\bf A fully implicit extended 3D MHD solver}}

MS K717 \\ Los Alamos National Laboratory \\ Los Alamos \\ NM 87545
\\
{\tt chacon@lanl.gov}\\
Dana A. Knoll\end{center}

We present results from our research on Jacobian-free Newton-Krylov
(JFNK) methods applied to the time-dependent, primitive-variable,
3D extended magnetohydrodynamics (MHD) equations. MHD is a fluid description
of the plasma state. While plasma is made up of independent (but coupled)
ion and electron species, the standard MHD description of a plasma
only includes ion time and length scales (one-fluid model). Extended
MHD (XMHD) includes nonideal effects such as nonlinear, anisotropic
transport and two-fluid (Hall and diamagnetic) effects. XMHD supports
so-called dispersive waves (whistler, ion acoustic), which feature
a quadratic dispersion relation $\omega \sim k^{2}$. In explicit
time integration methods, this results in a stringent CFL limit $\Delta
t_{CFL}\propto \Delta x^{2}$,
which severely limits their applicability to the study of long-frequency
phenomena in XMHD.

A fully implicit implementation promises efficiency (by removing the
CFL constraint) without sacrificing numerical accuracy \cite{knoll-jcp-03}.
However, the nonlinear nature of the XMHD system and the numerical
stiffness of its fast waves make this endeavor very difficult. Newton-Krylov
methods can meet the challenge provided suitable preconditioning is
available.

We propose a successful preconditioning strategy for the 3D primitive-variable
XMHD formalism. To focus the discussion, we consider the cold ion
limit ($T_{i}\ll T_{e}$) of a polytropic plasma ($p\propto n^{\gamma }$),
for which the XMHD model reads (in dimensionless units):\begin{eqnarray*}
\frac{\partial \rho }{\partial t} & + & \nabla \cdot (\rho \vec{v})=0,\\
\frac{\partial \vec{B}}{\partial t} & + & \nabla \times \vec{E}=0,\\
\frac{\partial (\rho \vec{v})}{\partial t}+\nabla \cdot \left[\rho
\vec{v}\vec{v}-\vec{B}\vec{B}\right. & - & \left.\rho \nu \nabla
\vec{v}+\overleftrightarrow{I}(p+\frac{B^{2}}{2})\right]=0,\\
\frac{\partial T_{e}}{\partial t}+\vec{v}\cdot \nabla T_{e} & + & (\gamma
-1)T_{e}\nabla \cdot \vec{v}=0,
\end{eqnarray*}
where $\vec{E}=-\vec{v}\times \vec{B}+\eta \nabla \times
\vec{B}-\frac{d_{i}}{\rho }(\vec{j}\times \vec{B}-\nabla p_{e})$
is the two-fluid generalized Ohm's law, $p=\rho T_{e}$ is the pressure,
and $\nu $, $\eta $ are the viscosity and resistivity, respectively.
The parameter $d_{i}=c/\omega _{pi}L$ is the so-called ion skin depth,
and is a measure of the importance of electron physics effects. Resistive
MHD is recovered in the limit $d_{i}\rightarrow 0$. A novel second-order,
cell-centered, conservative finite-volume discretization has been
recently developed in \cite{chacon-cpc-04} for the XMHD system above,
and will be used in this work. It is suitable for general curvilinear
geometries, solenoidal in $\mathbf{B}$ and $\mathbf{J}$, numerically
non-dissipative, and linearly and nonlinearly stable.

The XMHD preconditioner approach is based on {}``physics-based''
ideas \cite{chacon-jcp-02,chacon-jcp-03}, in which a hyperbolic system
of equations (which is diagonally submissive for $\Delta t>\Delta t_{CFL}$
\cite{chacon-jcp-02}) is {}``parabolized'' to arrive to a diagonally
dominant approximation of the original system, which is multigrid-friendly.
The use of approximate multigrid (MG) techniques to invert the {}``parabolized''
operator is a crucial step in the effectiveness of the preconditioner
and the scalability of the overall algorithm. The parabolization procedure
can be properly generalized using the well-known Schur decomposition
of a 2$\times $2 block matrix:{\small \[
\left[\begin{array}{cc}
D_{1} & U\\
L & D_{2}\end{array}
\right]=\left[\begin{array}{cc}
I & UD_{2}^{-1}\\
0 & I\end{array}
\right]\left[\begin{array}{cc}
D_{1}-UD_{2}^{-1}L & 0\\
0 & D_{2}\end{array}
\right]\left[\begin{array}{cc}
I & 0\\
D_{2}^{-1}L & I\end{array}
\right].\]
}In the context of XMHD, the resulting Schur complement $D_{1}-UD_{2}^{-1}L$
is a system of PDE's that couples the three plasma velocity components,
and needs to be inverted in a coupled manner. Nevertheless, a system
MG treatment is still possible since, when properly discretized, the
XMHD Schur complement is block diagonally dominant by construction,
and block smoothing is effective.

In this presentation, we will discuss the derivation and validity
of the physics-based preconditioner for resistive MHD and its generalization
to XMHD, the connection with Schur complement analysis, and the system-MG
treatment of the associated systems. We will demonstrate the algorithm
using the GEM challenge configuration \cite{gem}. Grid convergence
studies will demonstrate that CPU time scales scale optimally as $\mathcal{O}(N)$,
where $N$ is the number of unknowns, and that the number of Krylov
iterations scales as $\mathcal{O}(N^{0})$. Time convergence studies
will demonstrate a favorable scaling with time step $\mathcal{O}(\Delta t^{\alpha })$,
with $\alpha <1.0$.

\begin{thebibliography}{1}
\bibitem{knoll-jcp-03}D. A. Knoll, L. Chac\'{o}n, L. G. Margolin, and V. A. Mousseau, J.
Comput. Phys. \textbf{185}, 583 (2003)
\bibitem{chacon-cpc-04} \vspace{-8pt} L. Chac\'{o}n, Comp. Phys. Comm., \textbf{163}, 143
(2004)
\bibitem{chacon-jcp-02} \vspace{-8pt} L. Chac\'{o}n, D. A. Knoll, and J.
M. Finn, J. Comput.
Phys., \textbf{178}, 15 (2002)
\bibitem{chacon-jcp-03} \vspace{-8pt} L. Chac\'{o}n and D. A. Knoll, J. Comput. Phys.,
\textbf{188}, 573 (2003)
\bibitem{gem} \vspace{-8pt} J. Birn et al., J. Geophys. Res., \textbf{106}, 3715
(2001)\end{thebibliography}


\end{document}
