\documentclass{report}
\usepackage{amsmath,amssymb}
\setlength{\parindent}{0mm}
\setlength{\parskip}{1em}
\begin{document}
\begin{center}
\rule{6in}{1pt} \
{\large Eric de Sturler \\
{\bf The Convergence of Krylov Subspaces Methods with Recycling}}

Department of Mathematics \\ 460 McBryde \\ Virginia Tech \\ Blacksburg \\ VA 24061-0123
\\
{\tt sturler@vt.edu}\\
Michael L. Parks\end{center}

Many problems in science and engineering require the solution of a long
sequence of linear systems, with small changes from one matrix to the
next but substantial changes over multiple systems. We are particularly
interested in cases where both the matrix and the right hand side change
and systems are not available simultaneously. Such sequences arise in
time-dependent iterations, nonlinear systems of equations and
optimization, (distributed) parameter identification
and inverse problems, and many other problems.

In recent papers [1,2,3] we have
proposed methods to recycle selected subspaces from the Krylov
spaces generated for previous linear systems to improve the
convergence of subsequent linear systems. In this presentation, we
discuss several important convergence issues:
\begin{itemize}
\item the convergence of Krylov methods that recycle approximate
solution spaces, approximate invariant subspaces, and other relevant spaces,

\item the relevant perturbation theory for the spaces mentioned above for
sequences of matrices arising in a range of applications,

\item how fast our proposed methods learn to adapt to a changing
problem.
\end{itemize}
We provide experimental results for a range of problems from
tomography, nonlinear mechanics, large-scale design optimization, and
statistical mechanics.

\noindent
{\bf References}
\begin{enumerate}
\item
Michael L. Parks, Eric de Sturler, Greg Mackey, Duane D. Johnson, and Spandan Maiti,
{\em Recycling Krylov Subspaces for Sequences of Linear Systems}, SIAM
Journal on Scientific Computing (accepted with minor revisions), 2006,
available as Tech. Report
UIUCDCS-R-2004-2421, March 2004, from http://www-faculty.cs.uiuc.edu/\~{}sturler ,

\item
Misha Kilmer and Eric de Sturler,
{\em Recycling Subspace Information for Diffuse Optical Tomography},
SIAM Journal on Scientific Computing (accepted for publication),
2006, available from
\newline http://www-faculty.cs.uiuc.edu/\~{}sturler ,

\item
Shun Wang, Eric de Sturler, and Glaucio H. Paulino,
{\em Large-Scale Topology Optimization using Preconditioned Krylov
Subspace Methods with Recycling},
International Journal for Numerical Methods in Engineering (submitted), 2006,
available as Technical Report UIUCDCS-R-2006-2678
from http://www-faculty.cs.uiuc.edu/\~{}sturler.
\end{enumerate}


\end{document}
