\documentclass{report}
\usepackage{amsmath,amssymb}
\setlength{\parindent}{0mm}
\setlength{\parskip}{1em}
\begin{document}
\begin{center}
\rule{6in}{1pt} \
{\large Marco, M. Donatelli \\
{\bf Regularization by multigrid-type algorithms}}

Università dell'Insubria \\ Dip Fisica e Matematica \\ Via Valleggio \\ 11 \\ 22100 Como (CO) - Italy
\\
{\tt marco.donatelli@uninsubria.it}\end{center}

\newcounter{mycount}
\newenvironment{mylist}
{\begin{list}{\alph{mycount})}{\usecounter{mycount}}} {\end{list}}

We consider the de-blurring problem of noisy and blurred images in the
case of space invariant point spread functions (PSFs). The use of
appropriate boundary conditions leads to linear systems with structured
coefficient matrices related to space invariant operators like Toeplitz,
circulants, trigonometric matrix algebras etc. We can obtain an effective
and fast solver by combining the algebraic multigrid described in
\cite{ADS} with the Tikhonov regularization (see \cite{MGMTyk}). A
theoretical proof of such behavior is provided in \cite{AD}. A completely
alternative proposal is to use the latter algebraic multigrid (which is
designed ad hoc for structured matrices) with the low-pass projectors
typical of the classical geometrical multigrid employed in a PDEs
context. Thus, using an appropriate smoother, we obtain an iterative
regularizing method (see \cite{mgmreg}). For a comprehensive description
of the previous multigrid techniques for image restoration refer to
\cite{Dphd}. The resulting iterative regularizing multigrid method
proposed and discussed in \cite{mgmreg} is based on:
\begin{enumerate}
\item projection in a subspace where it is easier to
distinguish between the signal and the noise,
\item application of an iterative regularizing method in the projected subspace.
\end{enumerate}
Therefore any iterative regularizing method like conjugate
gradient (CG), conjugate gradient for normal equation (CGNE),
Landweber etc., can be used as smoother in our multigrid
algorithm. The projector is chosen according to \cite{ADS} in
order to maintain the same algebraic structure at each recursion
level and having a low-pass filter property, which is very useful in
order to reduce the noise effects. In this way, we obtain a better
restored image with a flatter restoration error curve and also in less
time than the auxiliary method used as smoother.

Like any multigrid algorithm, the resulting technique is
parameterized in order to have more degrees of freedom: a simple
choice of the parameters allows to devise a powerful regularizing method
whose main features are the following:
\begin{mylist}
\item it is used with early stopping like any regularizing iterative
method and its cost per iteration is about $1/3$ of the cost of the
method used as smoother (CG, Landweber, CGNE);
\item it can be adapted to work with all the boundary conditions used in
literature (Dirichlet, periodic, Neumann or anti-reflective) since the
basic algebraic multigrid considered in \cite{ADS,AD} is an optimally
convergent method for any of the involved structures (Toeplitz,
circulant, cosine-algebra or sine-algebra) which naturally arise from the
chosen boundary conditions;
\item the minimal relative restoration error with respect to the true
image is significantly lower with regard to the method used as smoother
and the associated curve of the relative restoration errors with respect
to the iterations is ``flatter'' (therefore the quality of the
reconstruction is not critically dependent on the stopping iteration);
\item when it is applied to the system $A{\bf f}=\bf g$ the minimal
relative error is comparable with regard to all the best known techniques
for the normal equations $A^TA{\bf f}=A^T\bf g$, but in this case the
convergence is much faster;
\end{mylist}
As direct consequence of c), the choice of the exact iteration
where to stop is less critical than in other regularizing
iterative methods while, as a consequence of d), we can choose
multigrid procedures which are extremely more efficient than
classical techniques without losing accuracy in the restored
image. Several numerical experiments show the effectiveness of our
proposals. A Theoretical analysis of multigrid methods is usually a
difficult task and a first largely used approach considers a two grid
method. In the same way, to proving the regularizing properties of our
multigrid methods, we provide some estimations on the filter factor of
the two level strategy.

Finally, it can be easily (by using a simple projection at every
step) combined with nonnegativity constraints. Moreover we propose a
possible generalization where the multigrid regularization is applied as
a one-step method: now the only parameter to choose is the number of
recursive calls (it works, in some sense, like a threshold parameter).

Work partly joint with Stefano Serra Capizzano.

\begin{thebibliography}{biblio}
\bibitem{AD}
\textsc{A. Aric\`o and M. Donatelli},
\textit{A V-cycle Multigrid for multilevel matrix algebras: proof of
optimality}, Numer. Math., to appear.
\bibitem{ADS}
\textsc{A. Aric\`o, M. Donatelli, and S. Serra Capizzano},
\emph{V-cycle optimal convergence for certain (multilevel) structured matrices},
SIAM J. Matrix Anal. Appl., 26--1 (2004) pp. 186--214.
\bibitem{MGMTyk}
\textsc{M. Donatelli},
\emph{A Multigrid for image deblurring with Tikhonov regularization},
Numer. Linear Algebra Appl., 12 (2005), pp. 715--729.
\bibitem{Dphd}
{\sc M. Donatelli}. {\em Image Deconvolution and Multigrid Methods},
PhD Thesis in Applied and Computational Mathematics, Univ. of Milano, December 2005.
\bibitem{mgmreg}
\textsc{M. Donatelli and S. Serra Capizzano},
\emph{On the regularizing power of multigrid-type algorithms},
SIAM J. Sci. Comput., in press.
\end{thebibliography}


\end{document}
