\documentclass{report}
\usepackage{amsmath,amssymb}
\setlength{\parindent}{0mm}
\setlength{\parskip}{1em}
\begin{document}
\begin{center}
\rule{6in}{1pt} \
{\large Andrei Dr{\u a}g{\u a}nescu \\
{\bf An abstract method for extending two-level preconditioners to multilevel preconditioners of comparable quality}}

Sandia National Laboratories \\ Optimization/Uncertainty Estimation Department \\ P O Box 5800 \\ MS 1110 \\ Albuquerque \\ 87185-1110
\\
{\tt aidraga@sandia.gov}\end{center}

We present an abstract method for designing a multilevel preconditioner
given a two-level preconditioner for an operator with positive definite
symmetric part.

If we denote by $A_{\mathrm{fine}}$ and $A_{\mathrm{coarse}}$ two
discrete versions of a continuous operator, then a two-level
preconditioner $T_{\mathrm{fine}}$ for $A_{\mathrm{fine}}$ can be
described in general by a function $T_{\mathrm{fine}} = {\mathcal
F}(A_{\mathrm{coarse}}^{-1}, A_{\mathrm{fine}})$, where it is assumed
that the evaluation of ${\mathcal F}$ requires a level-independent number
of applications of $A_{\mathrm{fine}}$ and $k$ applications of
$A_{\mathrm{coarse}}^{-1}$ ($k=1$ or $2$). The natural extension to a
multilevel preconditioner, consisting in replacing in $T_{\mathrm{fine}}$
the call to $A_{\mathrm{coarse}}^{-1}$ with a recursive call to
${\mathcal F}$, is known to sometimes produce multilevel preconditioners
of lower quality (e.g., for certain types of inverse problems). Based on
the idea that inverting $A_{\mathrm{fine}}$ essentially means to solve
the nonlinear equation \mbox{$X^{-1} - A_{\mathrm{fine}}= 0$}, we define
our multigrid preconditioner to be the first Newton iterate of the map
$X\mapsto X^{-1} - A_{\mathrm{fine}}$ starting at the ``natural''
multilevel preconditioner. For $k=1$, the resulting algorithm has a
W-cycle structure, and differs only slightly from the textbook version of
the W-cycle. Moreover, the method guarantees that the resulting
preconditioner maintains the approximation quality of the initial
two-level preconditioner. The quality of approximation is measured using
a certain distance function, which determines the degree to which two
operators with positive definite symmetric parts are spectrally
equivalent. We apply this method to designing and analyzing a multigrid
preconditioner for a linear advection-diffusion-reaction equation.


{\small Sandia is a multiprogram laboratory operated by Sandia
Corporation, a Lockheed-Martin Company, for the United States Department
of Energy under Contract DE-AC04-94AL85000.}


\end{document}
