\documentclass{report}
\usepackage{amsmath,amssymb}
\setlength{\parindent}{0mm}
\setlength{\parskip}{1em}
\begin{document}
\begin{center}
\rule{6in}{1pt} \
{\large Rainer Fischer \\
{\bf Structured Matrices, Multigrid Methods, and the Helmholtz Equation}}

Technical University of Munich \\ Boltzmannstr 3 \\ 85748 Garching \\ Germany
\\
{\tt rainer.fischer@mytum.de}\\
Thomas Huckle\end{center}

We wish to present a different approach to the multigrid solution of the
Helmholtz equation with constant coefficients. It is primarily based on
certain classes of structured matrices and their strong correspondence to
generating functions.
Discretization of the Helmholtz equation with certain boundary conditions
results in structured linear systems which are associated with generating
functions.
Depending on the kind of boundary conditions, the discretized Helmholtz
equation is a linear system of Toeplitz, tau, circulant, or DCT-III
type. By solving these systems with normal equations, we have the
advantage that the corresponding generating functions are nonnegative,
although they have a whole curve of zeros.

The multigrid methods we develop are especially designed for structured
matrix classes, making heavy use of the associated generating functions.
Over the last ten years, a specific theory of multigrid methods has been
developed for structured matrices whose generating functions have
isolated zeros. It is based on the AMG approach and the convergence
theory of Ruge and St\"uben.
In this work, we extend some of these theoretical results to the case of
generating functions with whole zero curves, and apply these modified
multigrid methods to the solution of the Helmholtz equation. We propose
two different strategies how this can be done.
\begin{itemize}
\item The first strategy is based on the idea of representing the whole
zero curve on all grids. For a multigrid method based on the Galerkin
approach, we can prove optimal two-grid convergence, but such a method is
computationally too expensive. Therefore we propose a rediscretization
technique where the zero curve is approximated on each grid. This results
in fast convergence and in coarse grid matrices with the same banded
structure as the given matrix. The only disadvantage of this approach is
that zero curves become larger on coarser levels, and therefore the
number of grids is limited.
\item The second strategy consists of splitting the original problem into
a fixed number of coarse grid problems. Corresponding to a generating
function with isolated zeros, each of these problems locally represents
one part of the zero curve. Each coarse grid problem is solved with a
standard multigrid method. We combine this splitting technique with the
first strategy to construct a faster and more robust multigrid solver.
\end{itemize}
We wish to discuss the use of our multigrid strategies not only as a
solver, but also as a preconditioner for Krylov subspace methods.
Moreover, the methods can also be applied to anisotropic linear systems
whose generating functions have a whole zero curve.


\end{document}
