\documentclass{report}
\usepackage{amsmath,amssymb}
\setlength{\parindent}{0mm}
\setlength{\parskip}{1em}
\begin{document}
\begin{center}
\rule{6in}{1pt} \
{\large Andrew Knyazev \\
{\bf Block Locally Optimal Preconditioned Eigenvalue Xolvers (BLOPEX)}}

Department of Mathematics \\ University of Colorado at Denver P O Box 173364 \\ Campus Box 170 \\ Denver \\ CO 80217-3364
\\
{\tt andrew.knyazev@cudenver.edu}\\
Ilya Lashuk\\
Merico Argentati\\
	Evgueni Ovtchinnikov\end{center}

Block Locally Optimal Preconditioned Eigenvalue Xolvers (BLOPEX) is a
package, written in C, that at present includes only one eigenxolver,
Locally Optimal Block Preconditioned Conjugate Gradient Method (LOBPCG).
BLOPEX supports parallel computations through an abstract layer. BLOPEX
is incorporated in the HYPRE package from LLNL and is availabe as an
external block to the PETSc package from ANL as well as a stand-alone
serial library. Hypre and PETSc packages provide high quality multigrid
and domain decomposition preconditioning on parallel clusters with
distributed or shared memory.

The LOBPCG method, suggested and developed by Andrew Knyazev [1] in the
past decade, recently attracts an increasing attention as a potential
alternative to the shift-and-invert Lanczos and preconditioned Davidson
methods due to its simplicity, robustness and fast convergence. Several
MATLAB, C, C++ and FORTRAN implementations of the LOBPCG are developed by
different groups, e. g., for such applications areas as structured
mechanics and electronic structure calculations.

Main LOBPCG features: a matrix-free iterative method for computing
several extreme eigenpairs of symmetric positive generalized
eigenproblems; a user-defined preconditioner; robustness with respect to
random initial approximations, variable preconditioners, and
ill-conditioning of the stiffness matrix; apparently optimal convergence
speed. Numerical comparisons suggest that LOBPCG may be a genuine block
analog for eigenproblems of the standard preconditioned conjugate
gradient method for symmetric linear systems.

We present initial scalability results using BLOPEX with Hypre and PETSc
on one BlueGene/L box solving eigenvalue problems of record sizes.

[1] A.V. Knyazev, "Toward the Optimal Preconditioned Eigensolver: Locally
Optimal Block Preconditioned Conjugate Gradient Method." SIAM Journal on
Scientific Computing 23 (2001), no. 2, pp. 517-541.


\end{document}
