\documentclass{report}
\usepackage{amsmath,amssymb}
\setlength{\parindent}{0mm}
\setlength{\parskip}{1em}
\begin{document}
\begin{center}
\rule{6in}{1pt} \
{\large Giovanni Lapenta \\
{\bf A Preconditioned Newton-Krylov Strategy for Moving Mesh Adaptation}}

MS: K717 \\ Los Alamos National Laboratory \\ Los Alamos NM87545
\\
{\tt lapenta@lanl.gov}\\
Luis Chac\'on\end{center}

We propose a new approach to adaptive mesh motion based on solving a
coupled self-consistent system for the physics equations and for the
grid generation equations. The key aspect of the new method is the
solution of the non-linear coupled system with a preconditioned
Newton-Krylov method. The present work described the approach and
focuses on the preconditioning techniques.

Adaptive grids are becoming an ever more common tool for high
performance scientific computing. We focus here on the type of
adaptation achieved by moving a constant number of points according
to appropriate rules, an approach termed moving mesh adaptation
(MMA). The approach we consider here is based specifically on
retaining a finite volume approach but allowing the grid to evolve
in time according to a grid evolution equation obtained from
minimization principles. The approach originates from the seminal
papers by Brackbill and Saltzmann[1] and by Winslow[2].

In the present paper we consider the fundamental question in the application
of MMA. Is it worth the effort? The literature is very rich and considerable
results have been obtained in designing MMA approaches that provide
grids that can indeed present the desired properties. But the question
of whether once the adaptive grids are used the simulations are actually
more cost effective remains largely unanswered.

We have revisited the question and have reached the conclusion that
in order to obtain an effective MMA strategy, three ingredients need
to be considered.

First is the effective formulation of the moving grid equations. In
1D the problem is benign, as error minimization leads to error
equiditribution and to a rigorous and simple minimization procedure.
In 2D and 3D the problem is more challenging but we have derived an
effective approach based on the classic approach by
Brackbill-Saltzmann-Winslow[1,2]. The crucial ingredients of our
approach are the formulation of the physics equations in a
conservative form and of the formulation of the grid generation
equations using harmonic mapping [3,4]. The independent variables of
the physics equations are changed from the physical to the logical
space and the equations are rewritten in the logical space in a
fully conservative form [4].

Second is the solution algorithm for the MMA method. Here we bring
a new development. The moving mesh equations and the physics equations,
derived by discretizing the problem under investigation on a moving
grid, form a tightly coupled system of algebraic non-linear equations.
Traditionally, the coupling is broken, the physics and grid equations
being solved separately in a lagged time-splitting approach. Each
time step is composed of two alternating steps: the physics equations
are solved on the current grid, the grid equations are then solved
using new information from the solution of the physics equations.
However, in presence of sharp fronts or other moving features, breaking
such coupling can lead to grid lagging with respect of the physics
equations, with adaptation resulting behind rather than on the moving
feature.

We avoid breaking the coupling and solve the full non-linear set of
physics and grid equation using the preconditioned Newton-Krylov
(NK) approach [5].

Third ingredient in a cost-effective grid adaptation is an efficient
preconditioning technique. In 1D a simple block tridiagonal approach
works effectively [6]. Each set of equations, for physics and for
grid generation, is preconditioned with a tridiagonal matrix
obtained by numerically approximating the corresponding diagonals in
the Jacobian. In 2D, we rely on a multigrid preconditioning strategy
where a crucial innovation is how to coarsen the information
relative to the adaptivity in the harmonic grid generation equations
[4].

In the present paper we describe the approach followed and we report
a number of examples to illustrate the performance of the new
approach.


[1] J.~U. Brackbill, J.~S. Saltzman, Adaptive zoning for singular
problems in 2 dimensions, J. Comput. Phys. 46~(3) (1982) 342--368.

[2] A.~Winslow, Adaptive mesh zoning by the equipotential method,
Tech. Rep. UCID-19062, Lawrence Livemore Laboratory (1981).

[3] V.~D. Liseikin, Grid generation methods, Springer, Berlin, New
York, 1999.

[4] L.~Chac\'{o}n, G.~Lapenta, A fully implicit, nonlinear adaptive
grid strategy, J. Comput. Phys., 212 (2006) 703--717.

[5] C.~T. Kelley, Iterative methods for linear and nonlinear
equations, SIAM, Philadelphia, 1995.

[6] G.~Lapenta, L.~Chac\'{o}n, Cost-effectiveness of fully implicit
moving mesh adaptation: a practical investigation in 1D, J. Comput.
Phys., submitted.


\end{document}
