\documentclass{report}
\usepackage{amsmath,amssymb}
\setlength{\parindent}{0mm}
\setlength{\parskip}{1em}
\begin{document}
\begin{center}
\rule{6in}{1pt} \
{\large Barry Lee \\
{\bf A Novel Algebraic Multigrid-Based Approach for Solving Maxwell's Equations}}

CASC L-561 \\ LLNL \\ P O Box 808 \\ Livermore \\ CA \\ 94551-0909
\\
{\tt lee123@llnl.gov}\\
Charles Tong\end{center}

This talk presents a new algebraic multigrid-based method for solving the
curl-curl formulation of Maxwell's equations discretized with edge elements. The
ultimate goal of this approach is two-fold. The first is to produce a
multiple-coarsening multigrid method with two approximately decoupled
hierarchies branching off at the initial coarse level, one resolving
the divergence-free error and the other resolving the curl-free error, i.e.,
a multigrid method that couples only on the finest level and mimics
a Helmholtz decomposition on the coarse levels. The second consideration
is to produce the hierarchies using a non-agglomerate coarsening scheme. To
roughly attain this two-fold goal, this new approach constructs the
first coarse level using topological properties of the mesh. In particular,
a discrete orthogonal decomposition of the finest edges is constructed by
dividing them into two sets, those forming a minimum spanning tree and the complement
set forming the cotree. Since the cotree edges do not form
closed cycles, these edges cannot support ``complete" near-nullspace gradient
functions of the curl-curl Maxwell operator. Thus, partitioning the finest
level matrix using this tree/cotree decomposition, the cotree-cotree submatrix
does not have a large near-nullspace. Hence, a non-agglomerate algebraic
multigrid method (AMG) that can handle strong positive and negative
off-diagonal elements can be applied to this submatrix. This cotree operator
is related to the initial coarse-grid operator for the divergence-free
hierarchy. The curl-free hierarchy is generated by a nodal Poisson operator
obtained by restricting the Maxwell operator to the space of gradients.
Unfortunately, because the cotree operator itself is not the initial
coarse-grid operator for the divergence-free hierarchy, the
multiple-coarsening scheme composed of the cotree matrix and its coarsening,
and the nodal Poisson operator and its coarsening does not give an
overall efficient method. Algebraically, the tree/cotree coupling on the
finest level, which is accentuated through smooth divergence-free
error, is too strong to be handled sufficiently only on the finest level.
In this new approach, these couplings are handled using oblique/orthogonal
projections onto the space of discretely divergence-free vectors. In the
multigrid viewpoint, the initial coarsening from the target fine level
to the divergence-free subspace is obtained using these
oblique/orthogonal restriction/interpolation operators in the Galerkin
coarsening procedure. The resulting coarse grid operator can be preconditioned
with a product operator involving a cotree-cotree submatrix and a topological
matrix related to a discrete Poisson operator. The overall iteration is
then a multigrid cycle for a nodal Poisson operator (the curl-free branch)
coupled on the finest grid to a preconditioned Krylov iteration for the fine
grid Maxwell operator restricted to the subspace of discretely divergence-free
vectors. Numerical results are presented to verify the effectiveness
and difficulties of this new approach for solving the curl-curl formulation of
Maxwell's equations.


\end{document}
