\documentclass{report}
\usepackage{amsmath,amssymb}
\setlength{\parindent}{0mm}
\setlength{\parskip}{1em}
\begin{document}
\begin{center}
\rule{6in}{1pt} \
{\large Jia Liu \\
{\bf Preconditioning Techniques for the Navier-Stokes Equations in Rotation Form}}

Department of Mathematics and Computer Science \\ Emory University \\ Atlanta GA 30322
\\
{\tt jliu8@emory.edu}\end{center}

We consider preconditioned iterative methods applied to
discretizations of the Navier-Stokes equations in 2D and 3D bounded
domains. Both unsteady and steady flows are considered. The
equations are linearized by Picard iteration. We make use of the
rotation form of the momentum equations, which has several
advantages from the linear algebra point of view.


We focus on two classes of preconditioners for the resulting
nonsymmetric saddle point problems, namely, block triangular
preconditioners and some variants of the Hermitian/Skew-Hermitian
splitting (HSS) preconditioner. Both types of preconditioners have
comparable cost per iteration, and make use of (standard) fast
solvers for elliptic scalar PDEs (convection-diffusion and
Poisson-type).


We compare the performance of both types of preconditioners with
regard to the mesh size, the Reynolds number, the time step, and
other problem parameters. Our experiments indicate that fast
convergence independent of problem parameters is achieved in many
cases. We include comparing experiments for both the rotation form
and convection diffusion form of the Navier-Stokes equations the
nonlinear iteration.


This is joint work with Prof. Michele Benzi.


\end{document}
