\documentclass{report}
\usepackage{amsmath,amssymb}
\setlength{\parindent}{0mm}
\setlength{\parskip}{1em}
\begin{document}
\begin{center}
\rule{6in}{1pt} \
{\large Yousef Saad \\
{\bf MIQR: A Multilevel Incomplete QR Preconditioner for Large Sparse Least-Squares Problems}}

Yousef Saad \\ University of Minnesota \\ Department of Computer Science and Engineering \\ 4-196 EE/CSci Building \\ 200 Union Street S E \\ Minneapolis \\ MN 55455
\\
{\tt saad@cs.umn.edu}\\
Na Li\end{center}


We present a Multilevel Incomplete QR (MIQR)
factorization for solving large sparse
least-squares problems. The algorithm builds the
factorization by exploiting structural
orthogonality in general sparse matrices. At any
given step, the algorithm finds an independent set
of columns, i.e., a set of columns that have
orthogonal patterns. The other columns are then
block orthogonalized against columns of the
independent set and the process is repeated
recursively for a certain number of levels on
these remaining columns. The final level matrix
is processed with a standard QR or Incomplete QR
factorization. Dropping strategies are employed
throughout the levels in order to maintain a good
level of sparsity. A few improvements to this
basic scheme are explored. Among these is the
relaxation of the requirement of independent sets
of columns. Numerical tests are proposed which
compare this scheme with the standard incomplete
QR preconditioner, the robust incomplete
factorization (RIF) preconditioner, and ARMS (on
the normal equations).


\end{document}
