\documentclass{report}
\usepackage{amsmath,amssymb}
\setlength{\parindent}{0mm}
\setlength{\parskip}{1em}
\begin{document}
\begin{center}
\rule{6in}{1pt} \
{\large Andreas Stathopoulos \\
{\bf PRIMME: PReconditioned Iterative MultiMethod Eigensolver. A robust, efficient and flexible hermitian eigenvalue software}}

Department of Computer Science \\ College of William and Mary \\ Williamsburg \\ VA 23187-8795
\\
{\tt andreas@cs.wm.edu}\\
James McCombs\end{center}

\small
The numerical solution of large, sparse, hermitian eigenvalue problems
continues to be one of the most computationally intensive tasks.
Iterative eigensolvers are routinely called to solve for a large number
of eigenvalues of matrices with dimension of excess of a million.
As with linear systems of equations, large dimensions almost always
necessitate the use of preconditioning. In addition, eigensolvers must
also deal with the issue of storing and orthogonalizing an increasingly
large set of eigenvectors.

In recent years, many preconditioned eigensolvers have emerged that
perform well in many, but not all applications. Notable examples are the
many variants of the Jacobi-Davidson method (JDQR, JDCG, JDQMR), the
Generalized-Davidson+1 method, and the LOBPCG method, while variations
of the traditional RQI and Inverse Iteration methods are still in use.
Until recently, however, general purpose software that implements these
methods both efficiently and robustly has been very scarce.
As a result, application developers
could not compare the various algorithms to find the most suitable for
their application and computing environment. Often, they would develop
in-house implementations of algorithms specifically tuned for their
application. Such in-house approaches, however, cannot benefit from
current advances in eigenvalue research. Such flexibility in choosing
from a variety of methods is missing in todays software.

In some cases of existing software, robustness has taken a secondary
role behind efficiency. Ideally, an eigensolver should find all the
required eigenvalues, in the shortest possible time, and produce an
orthonormal basis for their invariant space. It is well known, that
iterative eigensolvers cannot guarantee that eigenvalues are not missed.
However, certain algorithmic techniques can increase the confidence in
the computed results, albeit at a higher computational cost. This
increased confidence is sometimes needed in applications.

Although some of the above eigenvalue methods have been shown to provide
nearly optimal convergence for one eigenvalue in terms of matrix-vector
operations, the question of efficiency is a much more complex issue.
Actual execution times depend on the computing platform (hardware,
compiler, libraries), on the number of eigenvalues required, on the
quality of the preconditioner, and often on the problem solved.
Block methods have become a prerequisite for good cache performance, yet
maintaining good convergence with a large block size is not straightforward.
In addition, there is a multitude of techniques for restarting, locking,
stopping inner-outer iterations, that when properly implemented can
significantly improve the efficiency of the eigensolvers.

In view of the above, our group has developed a robust, efficient multi-method
software called PRIMME. PRIMME is based on a Davidson-type main iteration,
but it implements various techniques such block, locking, various projections
for preconditioning (e.g., Olsen's, Jacobi-Davidson, etc), CG-type restarting
(giving rise to JD+1, and LOBPCG-type methods), and adaptive inner-outer
iterations (allowing for JDQMR/JDCG or inexact Inverse Iteration type methods).
The implementation of all these techniques on top of a common platform,
allows PRIMME to transform to any of the above state-of-the-art eigenvalue
methods, as well as to hybrids representing arbitrary combinations of
techniques.

More than thirty features are controllable and tunable by the user.
However, this flexibility is not at the expense of usability. A complete set
of defaults is provided, and the user can simply select from a list
of twelve predefined methods. Alternatively, a method selection can be
further tuned by resetting particular features.
Such a multi-layer transparency addresses the different levels of expertise
of potential users, from end-users to eigenvalue experts.

Robustness decisions are prevalent in various implementation details of
PRIMME, such as multiple levels of convergence checking, validation processes,
out of order convergence of required eigenvalues, especially interior
eigenvalues, and consideration of numerical error. In cases where a choice
between robustness and efficiency has to be made, PRIMME favors robustness
by default, but it still implements the efficient approach as a user-defined
alternative.

Finally, the software runs both on parallel and sequential machines and can use
the optimized BLAS and LAPACK libraries of the target machines. Additionally,
the structure of PRIMME allows for a runtime capability of sensing both
the computing environment and the problem solved, and adapting the choice
of parameters accordingly. The potential of such a fully dynamic multi-method
is a particular focus of our current research.

In this talk, we present an overview of PRIMME and its interface,
and show some sample numerical results that demonstrate its robustness
and efficiency. In particular, we show that JDQMR and GD+1, two of the
supported methods, provide minimal execution time and number of iterations
respectively to a number of applications. The hope is that the choice
between the two can be fully automated within PRIMME.
Finally, we show a surprising result that for cases where a preconditioner
is not available, block JDQMR can be substantially better than ARPACK,
even for large numbers of eigenvalues.

The software is freely available under the lesser GPL license at:
\begin{center}
\tt{http://www.cs.wm.edu/$\sim$andreas/software/}
\end{center}


\end{document}
