\documentclass{report}
\usepackage{amsmath,amssymb}
\setlength{\parindent}{0mm}
\setlength{\parskip}{1em}
\begin{document}
\begin{center}
\rule{6in}{1pt} \
{\large Jari Toivanen \\
{\bf A Fast Iterative Solver for Acoustic Scattering by Objects in Layered Media}}

Center for Research in Scientific Computation \\ Box 8205 \\ North Carolina State University \\ Raleigh \\ NC 27695-8205
\\
{\tt jatoivan@ncsu.edu}\\
Kazufumi Ito\end{center}

We consider the computation of time-harmonic acoustic scattering
by objects in layered media. An example of such problem is the
scattering by a mine buried in sediment. The computational domain
can be tens or hundreds of meters long while the target requires
modeling of details smaller than one centimeter. A discretized
problem can have several billion degrees of freedom.

We decompose the computational domain into far-field and near-field
domains and then we perform a finite element discretization.
For the vastly larger far-field domain we use an orthogonal mesh
and a preconditioner based on a fast direct solver. For the
near-field domain a more standard preconditioner can be used.
The combination of these two defines a preconditioner for the
GMRES method. An essential implementation detail is the reduction
of iterations into a small sparse subspace. Due to this the memory
usage is essentially reduced and the GMRES method can be used
without restarts.

We present numerical results with two-dimensional and three-dimensional
problems demonstrating the efficiency of the proposed approach.
For example, we show that it is possible to solve problems with more
than a billion degrees of freedom on a modern PC.


\end{document}
