\documentclass{report}
\usepackage{amsmath,amssymb}
\setlength{\parindent}{0mm}
\setlength{\parskip}{1em}
\begin{document}
\begin{center}
\rule{6in}{1pt} \
{\large Rakhim Aitbayev \\
{\bf Multilevel Preconditioners for Nonselfadjoint or Indefinite Orthogonal Spline Collocation Problems}}

Department of Mathematics \\ New Mexico Tech \\ Socorro \\ NM 87801
\\
{\tt aitbayev@nmt.edu}\\
Bernard Bialecki\end{center}

\newtheorem*{theorem}{Theorem}\newcommand{\GP}{\mbox{${\cal G}$}}\newcommand{\SH}{\mbox{$\scriptscriptstyle H^{2}(\Omega)$}}\newcommand{\SL}{\mbox{$\scriptscriptstyle L^{2}(\Omega)$}}We develop and study symmetric multilevel preconditioners for the computation ofthe orthogonal spline collocation (OSC) solution of a Dirichlet boundary value problem(BVP) with a nonselfadjoint or an indefinite operator.The OSC solution is sought in the space of piecewise Hermite bicubic spline functionsdefined on a uniform partition.We consider an additive and a multiplicative multilevel preconditionersthat are used with the preconditioned conjugate gradient (PCG) method.Our results and algorithms are closely related to those in\cite{Aitbayev_Bialecki_2003},
\cite{Bialecki_1998},
\cite{Bialecki_Dryja_1997},and \cite{SBG_1996}.Let $\Omega$ be a unit square $(0,1)\times(0,1)$ with the boundary $\partial\Omega$,
andlet $x=(x_1,x_2)$.
We consider a BVP\begin{equation}\label{eq:bvp}Lu\equiv\sum_{i,j=1}^{2} a_{ij}(x)u_{x_ix_j}+\sum_{i=1}^{2} b_i(x) u_{x_i}+c(x)u = f(x),\;\;x\in\Omega,\quad u=0 \;\;\mbox{on}\;\;\partial\Omega.\end{equation}Operator $L$ could be non-selfadjoint or indefinite in $L^2$ inner product.We assume that the principal part of $L$ satisfies the uniform ellipticity conditionand that BVP (\ref{eq:bvp}) has a unique solution in $H^2(\Omega)$.Let $\pi_0$ be a uniform coarsest rectangular partition of $\Omega$.
We obtain a set of partitions$\{\pi_k\}_{k=0}^{K}$ by standard coarsening,
andlet $V_0\subset V_1\subset\ldots\subset V_K\equiv V_h$ be the set of corresponding nested spaces ofpiecewise Hermite bicubics that vanish on $\partial\Omega$.Let $\sum$ denote the two-dimensional composite Gauss quadrature corresponding to partition$\pi_h$ with 4 nodes in each element.
Let $\GP_h$ denote the corresponding set of Gauss points.The OSC discretization of BVP (\ref{eq:bvp}) is defined by\begin{equation}\label{eq:osc}u_h\in V_h,\quad Lu_h(\xi)=f(\xi),\quad\xi\in\GP_h,\end{equation}and it can be written as the operator equation $L_hu_h=f_h$ in the Hilbert space $V_h$ with theinner product $(v,w)_h=\sum vw$.We define and study multilevel additive$B_{\textup{\scriptsize a}}$ and multiplicative $B_{\textup{\scriptsize m}}$ preconditionersfor solving the normal equation $L_h^*L_hu_h=L_h^*f_h$,
where $L_h^*$ is the adjoint to $L_h$.The implementation of $B_{\textup{\scriptsize a}}$ and $B_{\textup{\scriptsize m}}$is based on relationships between basis functions for two consecutive partitionsand the implementation of $B_{\textup{\scriptsize m}}$ is similar to that forV(1,1)-cycle with the Gauss-Seidel smoothing.A problem on the coarsest partition is assumed sufficiently small,
and it is solved exactly.The computational cost of the preconditioning algorithms is $\mbox{O}(N_K)$.The following is our main result.\begin{theorem}There are positive independent of $h$ and $K$ constants$\alpha_{\textup{\scriptsize a}}$,$\beta_{\textup{\scriptsize a}}$,$\alpha_{\textup{\scriptsize m}}$,and$\beta_{\textup{\scriptsize m}}$,such that\begin{eqnarray}\label{eq:spectral_equiv}&&\alpha_{\textup{\scriptsize a}}\,
(B_{\textup{\scriptsize a}}v,v)_h\leq (L_h^*L_hv,v)_h \leq\beta_{\textup{\scriptsize a}}\,
(B_{\textup{\scriptsize a}}v,v)_h,\quad v\in V_h,\\\nonumber&&\alpha_{\textup{\scriptsize m}}\,
(B_{\textup{\scriptsize m}}v,v)_h\leq (L_h^*L_hv,v)_h \leq\beta_{\textup{\scriptsize m}}\,
(B_{\textup{\scriptsize m}}v,v)_h,\quad v\in V_h.\end{eqnarray}\end{theorem}We present numerical results that demonstrate the efficiency ofour preconditioning algorithms.\begin{thebibliography}{1}\bibitem{Aitbayev_Bialecki_2003}{\sc R.~Aitbayev and B.~Bialecki},
{\em A preconditioned conjugate gradientmethod for nonselfadjoint or indefinite orthogonal spline collocationproblems},
SIAM J.
Numer.
Anal.,
41 (2003),
pp.~589--604.\bibitem{Bialecki_1998}{\sc B.~Bialecki},
{\em Convergence analysis of orthogonal spline collocationfor elliptic boundary value problems},
SIAM J.
Numer.
Anal.,
35 (1998),pp.~617--631.\bibitem{Bialecki_Dryja_1997}{\sc B.~Bialecki and M.~Dryja},
{\em Multilevel additive and multiplicativemethods for orthogonal spline collocation problems},
Numer.
Math.,
77 (1997),pp.~35--58.\bibitem{SBG_1996}{\sc B.~F.
Smith,
P.~E.
Bj{\o}rstad,
and W.~D.
Gropp},
{\em Domain Decomposition:Parallel Multilevel Methods for Elliptic Partial Differential Equations},Cambridge University Press,
New York,
1996.\end{thebibliography}

\end{document}
