\documentclass{report}
\usepackage{amsmath,amssymb}
\setlength{\parindent}{0mm}
\setlength{\parskip}{1em}
\begin{document}
\begin{center}
\rule{6in}{1pt} \
{\large Erik B{\"a}ngtsson \\
{\bf Numerical simulation of glacial rebound using preconditioned iterative solution methods}}

Department of Information Technology,Box 337,SE-751 05 Uppsala,Sweden
\\
{\tt erikba@it.uu.se}\\
Maya Neytcheva\end{center}

We consider the problem to compute the stress ($\sigma$) and
displacement ($\bf{u}$) fields in a (visco)elastic inhomogeneous layered
media, in response to a surface load.
The underlying physical phenomenon, which is modeled, is glacial advance
and recession, and the post-glacial rebound caused by the latter, which
reflects the viscoelastic properties of the mantle. The nowadays observed
residual rebound
implies that the lithosphere might not completely reach equilibrium
within itself before another glacial period.
Therefore post-glacial rebound has to be taken into account
in the context of other problems, such as for example predicting safety
of nuclear waste repositories.

The material incremental momentum equation for quasi-static infinitesimal
perturbations of a stratified, compressible fluid Earth, initially in
hydrostatic equilibrium, subject to gravitational forces but neglecting
internal forces (cf. $[5]$) is
\begin{displaymath}
\underset{(A)}{\underbrace{\nabla\cdot\sigma}} +
\underset{(B)}{\underbrace{\nabla(\bf{u}\cdot\nabla p^{(0)})}} +
\underset{(C)}{\underbrace{\rho^{(\Delta)}\bf{g}^{(0)}}} +
\underset{(D)}{\underbrace{\rho^{(0)}\nabla\bf{g}^{(\Delta)}}} = \bf{0}.
\end{displaymath}
Here, term $(A)$ describes the force from spatial gradients in stress. If
a large elastic solid is put in a gravitational field, it becomes
gravitationally pre-stressed with pressure $p^{(0)}$. This pressure can
be regarded as an initial condition imposed on the problem and does not
cause deformations. Term $(B)$ represents the advection of
this pre-stress and describes how it is carried by the moving
material. Terms $(C)$ and $(D)$ describe perturbations of the
gravitational force and gravitational acceleration due to changes of
density.

In the present study, an incompressible non-selfgravitating (flat) Earth
model is used, which implies constant gravity field and constant density,
so that these two terms vanish. Term (B)
is further simplified assuming that the advection term describes the
advection in the direction of the gravity field only.

Incorporating the above simplifications with respect to terms
$(B)$, $(C)$ and $(D)$, we obtain the following form of the governing
equilibrium equation
\begin{equation}
\begin{array}{l}
\nabla\cdot \sigma +
\rho^{(0)}g^{(0)} \nabla(u_d) = \mathbf{0} \quad
\mathbf{x}\in\Omega\subset\mathbb{R}^d, d=2,3
\end{array}
\label{eq_rebound2}
\end{equation}
with suitable boundary conditions.

In its full complexity, the model includes viscoelastic constitutive
relations. In this work we discuss a purely elastic material behavior
only, as is analyzed in $[5]$, for instance.

Problem (\ref{eq_rebound2}) is discretized using stable mixed finite
element pairs or a suitable stabilized formulation,
which lead to a system of linear equations with a nonsymmetric
two-by-two block matrix of a saddle point form.

Results from numerical experiments solving the so-arising algebraic
system with preconditioned iterative solution methods are presented.
Several preconditioning strategies are tested, based on the techniques
and experience described in $[4]$,
$[3]$, $[1]$, $[2]$ and other authors.
The performance of the tested preconditioned iterative solution methods
is compared with that of a commercial FEM package solver.

\begin{table}[htb]
\begin{tabular}{c p{12.5cm}}
$[1]$&Axelsson O, Neytcheva M.
Preconditioning methods for linear systems arising in constrained
optimization problems,
{\em Num. Lin. Alg. Appl.}, 10 (2003), 3-31.
\\
$[2]$ &
Axelsson O, Neytcheva M.
Preconditioning methods for for constrained optimization problems
with applications for the linear elasticity equations
Report 0302, January 2003, Department of Mathematics,
University of Nijmegen, The Netherlands.
\\
$[3]$ &
Elman H, Silvester D, Wathen A. Performance and analysis of saddle
point preconditioners for the discrete steady-state Navier-Stokes equations.
\textit{Num. Mat.}, {\bf 90} (2002) 665-688.
\\
$[4]$& Klawonn A. Block-triangular preconditioners for saddle point
problems with a penalty term. {\em SIAM J. Sci. Comp.} 1998;
19:172-184.
\\
$[5]$&
Klemann V, Wu P, Wolf D. Compressible viscoelasticity: stability of
solutions for homogeneous plane-Earth models,
\textit{Geophys. J.}, 2003; \textbf{153}:569-585.\\
\end{tabular}

\end{table}


\end{document}
