\documentclass{report}
\usepackage{amsmath,amssymb}
\setlength{\parindent}{0mm}
\setlength{\parskip}{1em}
\begin{document}
\begin{center}
\rule{6in}{1pt} \
{\large Matthias Bollhoefer \\
{\bf Inverse-Based Multilevel ILU Preconditioners --- Theory, Applications and Software}}

Inst. of Mathematics \\ MA 4-5 \\ TU Berlin \\ Str. des 17. Juni 136 \\ 10623 Berlin \\ GERMANY
\\
{\tt bolle@math.tu-berlin.de}\\
Yousef Saad\end{center}

We will present multilevel ILU strategies for large scale
problems that are mainly based
on keeping the inverse triangular factors bounded.
This approach is based on combining several single components
that finally form the multilevel algorithm.

\begin{enumerate}
\item
The components cover reordering strategies like RCM, independent
sets or a more recent strategy ddPQ \cite{Saa03a}, which partially reorders
the matrix with respect to diagonal dominance and fill.
\item
In the ILU part of the algorithm, the preprocessed matrix is partially
factored using
an inverse based ILU \cite{Bol03a,LiSC03} with diagonal
pivoting that keeps the inverse triangular
factors bounded. From the theoretical point of view this
can be interpreted as keeping the approximate
triangular factors $L$ and $U$
and their inverses close to each other.
\item The combination of the first two parts leads to a multilevel
algorithm that repeatedly uses these templates.
\end{enumerate}

The algorithm has been implemented in a new software package that includes
treatment of general real and complex matrices as well
as SPD matrices.
The effectiveness of this approach using the package is demonstrated.

\begin{thebibliography}{1}

\bibitem{Bol03a}
M.~Bollh\"ofer.
\newblock A robust and efficient {I}{L}{U} that incorporates the growth of the
inverse triangular factors.
\newblock {\em {SIAM} J. Sci. Comput.}, 25(1):86--103, 2004.

\bibitem{LiSC03}
N.~Li, Y.~Saad, and E.~Chow.
\newblock Crout versions of {I}{L}{U} for general sparse matrices.
\newblock {\em {SIAM} J. Sci. Comput.}, 25(2):716--728, 2004.

\bibitem{Saa03a}
Y.~Saad.
\newblock Complete pivoting ilu: a multilevel approach.
\newblock Research Report UMSI 2003-191, University of Minnesota, Super
Computing Institute, Minneapolis, Minnesota, 2003.

\end{thebibliography}


\end{document}
