\documentclass{report}
\usepackage{amsmath,amssymb}
\setlength{\parindent}{0mm}
\setlength{\parskip}{1em}
\begin{document}
\begin{center}
\rule{6in}{1pt} \
{\large Efstratios Gallopoulos \\
{\bf Computating pseudospectra of parameter-dependent matrices}}

Department of Computer Engineering & Informatics \\ University of Patras \\ 26500 Patras \\ Greece
\\
{\tt stratis@ceid.upatras.gr}\\
Costas Bekas\\
Kyriakos Petrakos\end{center}

The motivation for studying the matrix pseudospectrum is that it
frequently provides information regarding matrix modeled processes
that is not readily available from the spectrum. Only a few years
ago, computing pseudospectra was considered to be very expensive
even for moderately sized matrices. Recent research, however, has
brought down the cost considerably and makes their computation
feasible even for relatively large matrices. A general assumption
in these efforts was that the matrix under study does not change.
In this presentation we relax this constraint and consider the
pseudospectrum of matrices that depend on parameters, such as
time. Typical applications in which such matrices arise include
the numerical solution of PDEs with time varying coefficients and
bifurcation problems in dynamical systems using Newton-Krylov
schemes. The objective is to design methods that follow the
evolution, with $t$, of the pseudospectrum
$\Lambda_\epsilon(A(t))$ of some matrix function $A(t)$. A
recently designed method for the parallel computation of
pseudospectra, called {\sc psdm}, will serve as a primary tool
of these investigations.

Work conducted in the context of and supported in part by a
University of Patras CARATHEODORY grant.


\end{document}
