\documentclass{report}
\usepackage{amsmath,amssymb}
\setlength{\parindent}{0mm}
\setlength{\parskip}{1em}
\begin{document}
\begin{center}
\rule{6in}{1pt} \
{\large Olivier, P. Le Maitre \\
{\bf Multigrid Method for Spectral Propagation of Uncertainty in Diffusion Equations}}

Universite d'Evry Val d'Essonne \\ IUP \\ 40 rue du Pelvoux \\ CE1455 \\ 91 020 Evry Courcourone cedex \\ France
\\
{\tt olm@cemif.univ-evry.fr}\\
Bert, J. Debusschere\\
Roger, G. Ghanem\\
	Knio, Najm\end{center}

Stochastic modeling of two-dimensional steady and unsteady
diffusion equations in presence of uncertainties in the
diffusivity field, forcing term and/or boundary conditions
is considered, using spectral expansions. Diffusivity and
source fields, as well as the boundary conditions, are
stochastic processes which are first represented using
their respective Karhunen-Lo\`eve (KL) expansions. When
truncated, the KL expansions involve a finite set of independent
random variables that multiply the eigen-modes of the
correlation kernels of the corresponding processes.
Propagation and quantification of the impact of these
uncertainties on the solution is performed by means of a
Polynomial Chaos (PC) expansion, whose spectral coefficients are
determined using a Galerkin approach. The Galerkin projection
leads to a set of coupled problems for the PC
coefficients of the solution which are functions of space and
time. Spatial finite-difference discretization of these coupled
problems results in a large system of equations, whose dimension
necessitates the use of iterative approaches in order to obtain
the solution within a reasonable computational time.
To accelerate the convergence of the iterative technique, a
multigrid method, based on spatial coarsening, is implemented.
Numerical experiments show good scaling properties of
the method, both with respect to the number of spatial grid points
and the stochastic resolution level.


\end{document}
