\documentclass{report}
\usepackage{amsmath,amssymb}
\setlength{\parindent}{0mm}
\setlength{\parskip}{1em}
\begin{document}
\begin{center}
\rule{6in}{1pt} \
{\large Heidi K. Thornquist \\
{\bf Fixed-Polynomial Approximate Spectral Transformations for Preconditioning the Eigenvalue Problem}}

P.O. Box 5800 \\ MS 1110 \\ Albuquerque \\ NM  87185-1110
\\
{\tt hkthorn@sandia.gov}\\
Danny C. Sorensen\end{center}

Arnoldi's method is often used to compute the eigenvalues
and eigenvectors of large, sparse matrices. Various
techniques can by employed to improve the convergence of
this method, some resulting in considerable expense.
We will present an efficient preconditioning scheme that
accelerates the convergence of Arnoldi's method to the
eigenvalues of smallest magnitude for Hermitian and
non-Hermitian matrices. This scheme uses a fixed-polynomial
operator to approximate the spectral transformation.
When used with the Implicitly Restarted Arnoldi (IRA) method,
this achieves the effect of a spectral transformation without
factoring a matrix.

Numerical results indicate that this employing this
preconditioning scheme with IRA is very effective. It
is just as accurate as IRA in regular mode, more accurate
than using preconditioned GMRES started anew for each
linear solve, and considerably less expensive than either.
Furthermore, using a fixed-polynomial operator with IRA
compares favorably with two other current methods:
Jacobi-Davidson and Locally Optimal Block Preconditioned
Conjugate Gradient.


\end{document}
