\documentclass{report}
\usepackage{amsmath,amssymb}

\def\mathbi#1{\textbf{\em #1}}
\def\BA{\bf{A}}
\def\BB{\bf{B}}
\def\bn{\bf{n}}
\def\CP{\mathcal{P}}
\def\CV{\mathcal{V}}
\def\CI{\mathcal{I}}

\newcommand{\gradt}{\nabla\cdot}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\Cplex}{\mathbb{C}}

\newcommand{\Vecc}[2]{ \left(
	\begin{array}{c}
	#1 \\ #2
	\end{array}
	\right) }

\newcommand{\Matrr}[4]{ \left(
	\begin{array}{cc}
	#1 & #2 \\
	#3 & #4
	\end{array}
	\right) }

\newcommand{\calB}{{\mathcal B}}
\newcommand{\calBB}{{\mathcal B}^{\Box}}
\newcommand{\calK}{\mathcal{K}}
\newcommand{\bfA}{{\mathbf A}}
\newcommand{\bfb}{{\mathbf b}}
\newcommand{\bfr}{{\mathbf r}}
\newcommand{\bfx}{{\mathbf x}}
\newcommand{\bfxex}{{\mathbf x}_{\mbox{\scriptsize $\star$}}}
\newcommand{\Dim}{\mathop{\mathrm{dim\ }}}

\begin{document}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{center}
{\large
{\bf Preconditioning techniques for the Navier-Stokes equations \\
	in rotation form}}

	Jia Liu \\
	Dept.~of Mathematics and Computer Science \\
	Emory University, Atlanta GA 30322 \\
	{\tt jliu8@emory.edu} \\
	Michele Benzi
\end{center}
We consider preconditioned iterative methods applied to
discretizations of the Navier-Stokes equations in 2D and 3D
bounded domains. Both unsteady and steady flows are
considered. The equations are linearized by Picard
iteration. We make use of the rotation form of the momentum
equations, which has several advantages from the linear
algebra point of view.

We focus on two classes of
preconditioners for the resulting nonsymmetric saddle point
problems, namely, block triangular preconditioners and some
variants of the Hermitian/Skew-Hermitian splitting (HSS)
preconditioner. Both types of preconditioners have
comparable cost per iteration, and make use of (standard)
fast solvers for elliptic scalar PDEs (convection-diffusion
and Poisson-type).

We compare the performance of both
types of preconditioners with regard to the mesh size, the
Reynolds number, the time step, and other problem
parameters. Our experiments indicate that fast convergence
independent of problem parameters is achieved in many cases.
We include comparing experiments for both the rotation form
and convection diffusion form of the Navier-Stokes equations
the nonlinear iteration.



	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\end{document}

