\documentclass{report}
\usepackage{amsmath,amssymb}

\def\mathbi#1{\textbf{\em #1}}
\def\BA{\bf{A}}
\def\BB{\bf{B}}
\def\bn{\bf{n}}
\def\CP{\mathcal{P}}
\def\CV{\mathcal{V}}
\def\CI{\mathcal{I}}

\newcommand{\gradt}{\nabla\cdot}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\Cplex}{\mathbb{C}}

\newcommand{\Vecc}[2]{ \left(
	\begin{array}{c}
	#1 \\ #2
	\end{array}
	\right) }

\newcommand{\Matrr}[4]{ \left(
	\begin{array}{cc}
	#1 & #2 \\
	#3 & #4
	\end{array}
	\right) }

\newcommand{\calB}{{\mathcal B}}
\newcommand{\calBB}{{\mathcal B}^{\Box}}
\newcommand{\calK}{\mathcal{K}}
\newcommand{\bfA}{{\mathbf A}}
\newcommand{\bfb}{{\mathbf b}}
\newcommand{\bfr}{{\mathbf r}}
\newcommand{\bfx}{{\mathbf x}}
\newcommand{\bfxex}{{\mathbf x}_{\mbox{\scriptsize $\star$}}}
\newcommand{\Dim}{\mathop{\mathrm{dim\ }}}

\begin{document}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{center}
{\large
{\bf Preconditioning analysis for weighted
Toeplitz regularized least squares problems}}

	Lung Chak Chan \\
	Dept. of Mathematics, The University of Hong Kong \\
	Pokfulam Road, Hong Kong \\
	{\tt lungchak@graduate.hku.hk}
\end{center}

We consider
preconditioners for weighted
Toeplitz regularized least squares (WTRLS) problems
$$
\min_z \| A z - b \|_2^2
$$
where the rectangular coefficient matrix $A$ and the
right-hand side vector $b$ are of the form:
$$ 
A =
\left [
\begin{array}{c}
W_1 H \\
\mu W_2 L \\
\end{array}
\right ] \in \Reals^{(m_{1}+m_{2}) \times m_{1}}
\quad {\rm and}\quad  b =
\left [
\begin{array}{c}
W_1 f \\
 0 \\
\end{array}
\right ] \in \Reals^{m_{1}+m_{2}} \,.
$$
Here $H \in \Reals^{m_{1} \times m_{1}}$ is a Toeplitz matrix, $L
\in \Reals^{m_{2} \times m_{1}}$ is the first-order or second-order
difference matrix, $W_1 \in \Reals^{m_{1} \times m_{1}}$ and $W_2
\in \Reals^{m_{2} \times m_{2}}$ are non-scalar diagonal matrices
with real positive entries, $f \in \Reals^{m_{1}}$ is a given
right-hand side vector and $\mu >0$ is a regularization parameter.


Iterative methods using circulant preconditioners has been proposed 
since 1980's. 
The efficiency of solving systems are
greatly enhanced by these indirect methods.  Since then,
preconditioners based on circulant matrices were suggested for
solving the WTRLS problems.  However, most of them do not perform
satisfactorily and the results of convergence rates can be disappointing.


Benzi and Ng 
considered the above system
where $W_{2}L$ is replaced by $I$, and proposed a new approach to
solve it which is based on an augmented system formulation. We now
adopt a similar approach that transforms the problem into a
$3 \times 3$ block linear system.

The WTRLS problems turn out to be a quadratic minimization problem, which is equivalent to solving the linear system
$$ %\label{new}
(H^{T}W_{1}^{-2}H + \mu^{2} L^{T}W_{2}^{-2}L)z = H^{T} W_{1}^{-2} f~.
$$


The above linear system can overall be rewritten as the following $3 \times 3$ block system:
$$ 
\left[ \begin{array}{ccc} W_{1}^{-2} & 0 & H \\ 0 & W_{2}^{-2} & \mu L \\ H^{T} & \mu L^{T} & 0 \end{array} \right] \left[ \begin{array}{c} x \\ y \\ z \end{array} \right] = \left[ \begin{array}{c} 0 \\ 0 \\ -H^{T} W_{1}^{2} f \end{array} \right]~.
$$


There are not any particular results discussing how
to solve a general $3 \times 3$ system effectively.  
The four upper left blocks are considered as one single
block and the corresponding two upper right blocks
$[H^{T} \ \mu L^{T}]^{T}$ together
still attain full rank, so we treat this system as
a $2 \times 2$ block case 
and consider methods developed for solving indefinite systems.  
Here we shall adopt constraint
preconditioner 
and Hermitian skew-Hermitian splitting (HSS)
preconditioner 
and investigate them in details.
The formulation and actual preconditioning settings
for WTRLS problems will be discussed.
A test problem is performed to demonstrate the convergence behaviour
using these iterative methods.  We shall see that the number of iterations
mainly depend on the condition numbers of $W_{1}$ and $W_{2}$ as
well as some factors related to each preconditioners. 
In particular, this effect is more significant in
the case of using HSS preconditioner.  The iteration results and
the weaknesses for both preconditioners can be explained
with some analyses on the spectra for them.

\end{document}
