\documentclass{report}
\usepackage{amsmath,amssymb}
\setlength{\parindent}{0mm}
\setlength{\parskip}{1em}
\begin{document}
\begin{center}
\rule{6in}{1pt} \
{\large Matthias Bolten \\
{\bf Parallelization of a multigrid method for PDEs with open boundary conditions}}

Department of Mathematics and Science \\ University of Wuppertal \\ D-42097 Wuppertal \\ Germany
\\
{\tt bolten@math.uni-wuppertal.de}\end{center}

The solution of PDEs with so-called open boundary conditions plays an
important role in different areas, e.g.~in the simulation of molecules or
in aerospace engineering. An example is the Poisson equation
\begin{align*}
- \Delta u(x) & = f(x), \text{ for all } x \in \mathbb{R}^3, \text{ where}\\
u(x) & \stackrel{\| x \|_2 \rightarrow \infty}{\longrightarrow} 0,
\end{align*}
as it is needed in molecular dynamics simulation. In~\cite{art:BOLT07} we
extended the method developed by Washio and Oosterlee
in~\cite{art:WASH00}. In contrast to the original method the modification
guarantees that the number of levels depends on the grid size of the
finest level, only. The optimality of the method w.r.t.~the number of
unknowns was proven in~\cite{phd:BOLT08}.

For large-scale simulations, as they are carried out e.g.~in the chemical
industry, a highly scalable method is needed. The parallelization of the
method is not straightforward, as either an imbalance of the work occurs
on some levels, or a sophisticated communication scheme is required.

In the talk the multigrid method is described, the two parallelization
approaches are outlined and a comparison of the two is presented.

\begin{thebibliography}{1}

\bibitem{phd:BOLT08}
M.~Bolten.
\newblock {\em Multigrid methods for structured grids and their application in
particle simulation}.
\newblock PhD thesis, Bergische Universit{\"a}t Wuppertal, Wuppertal, 2008.

\bibitem{art:BOLT07}
M.~Bolten.
\newblock Hierarchical grid coarsening for the solution of the poisson equation
in free space.
\newblock {\em Electron. Trans. Numer. Anal.}, 29:70--80, 2007.

\bibitem{art:WASH00}
T.~Washio and C.~W. Oosterlee.
\newblock Error analysis for a potential problem on locally refined grids.
\newblock {\em Numer. Math.}, 86(3):539--563, 2000.

\end{thebibliography}


\end{document}
