\documentclass{report}
\usepackage{amsmath,amssymb}
\setlength{\parindent}{0mm}
\setlength{\parskip}{1em}
\begin{document}
\begin{center}
\rule{6in}{1pt} \
{\large Abdul Hanan Sheikh \\
{\bf A scalable Helmholtz solver combining the deflation with shifted Laplace preconditioner.}}

Mekelweg 4 \\ EWI Building \\ Delft University of Technology \\ 2628CD Delft \\ Netherlands
\\
{\tt hanangul12@yahoo.co.uk}\\
Domenico Lahaye\\
Kees Vuik\end{center}

Our object is to develop high performance iterative solution algorithm
for solving the discrete Helmholtz equation modeling wave propagation on
large scale. Ingredients in our work are the shifted Laplace
preconditioner and deflation. The development of the shifted Laplace
preconditioner for the Helmholtz equation was a breakthrough in the
development of efficient solution techniques for the Helmholtz equation.
The distinct feature of this preconditioner is the introduction of a
complex shift, effective introducing damping of wave propagation in the
approximate solve. This preconditioner was extensively discussed in
various texts and applied in a number of different contexts. Although
performant, the resulting algorithm is not truly scalable. The bigger the
wavenumber, the more spectrum scatters away from one, hampering the
convergence. Idea of projection has been used since long to deflate
unfavorable eigenvalues. By inducting eigenvectors corresponding to
unwanted eigenvalues, better convergence for CG and GMRES has been
reported in various texts. We also combine the idea of deflation with
shifted Laplace preconditioner, which leads to a scalable Helmholtz
solver, in the sense iterations does not depend upon parameters.
We provide a convergence analysis. We perform a Fourier two-grid analysis
of one-dimensional model problem with Dirichlet boundary conditions
discretized by a second order accurate finite difference scheme. The
components analyzed are the shifted Laplace preconditioner used as
smoother, full-weighting and linear interpolation intergrid transfer
operators, and a Galerkin coarsening scheme. This Fourier analysis
results in a closed form expression for the eigenvalues of the two-grid
operator. This expressions shows that the spectrum is favourable for
convergence of Krylov subspace methods.
We apply the deflated shifted Laplace preconditioner to two-dimensional
model problems method with constant and non-constant wave numbers and
Sommerfeld boundary conditions discretized by second order accurate
finite difference scheme on uniform meshes. Numerical results show that
the number of GMRES iterations is wave-number independent.



\end{document}
