\documentclass{report}
\usepackage{amsmath,amssymb}

\def\mathbi#1{\textbf{\em #1}}
\def\BA{\bf{A}}
\def\BB{\bf{B}}
\def\bn{\bf{n}}
\def\CP{\mathcal{P}}
\def\CV{\mathcal{V}}
\def\CI{\mathcal{I}}

\newcommand{\gradt}{\nabla\cdot}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\Cplex}{\mathbb{C}}

\newcommand{\Vecc}[2]{ \left(
	\begin{array}{c}
	#1 \\ #2
	\end{array}
	\right) }

\newcommand{\Matrr}[4]{ \left(
	\begin{array}{cc}
	#1 & #2 \\
	#3 & #4
	\end{array}
	\right) }

\newcommand{\calB}{{\mathcal B}}
\newcommand{\calBB}{{\mathcal B}^{\Box}}
\newcommand{\calK}{\mathcal{K}}
\newcommand{\bfA}{{\mathbf A}}
\newcommand{\bfb}{{\mathbf b}}
\newcommand{\bfr}{{\mathbf r}}
\newcommand{\bfx}{{\mathbf x}}
\newcommand{\bfxex}{{\mathbf x}_{\mbox{\scriptsize $\star$}}}
\newcommand{\Dim}{\mathop{\mathrm{dim\ }}}

\begin{document}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{center}
{\large
{\bf A greedy strategy for coarse-grid selection}}

	Scott MacLachlan \\
	Dept.~of Computer Science and Engineering \\
	University of Minnesota, 4-192, EE/CS Building \\
	200 Union Street SE, Minneapolis MN 55455 \\
	{\tt maclach@cs.umn.edu} \\
	Yousef Saad
\end{center}
In recent years, substantial effort has been focused on
developing methods capable of solving the large linear
systems that arise from the discretization of partial
differential equations. This research has been driven by
application scientists' demands of higher accuracy, in both
mathematical modeling and computational solution. Efficient
solution of many of these linear systems is possible only
through the use of multilevel solution techniques. While
highly optimized algorithms may be developed using knowledge
about the origins of the matrix problem to be considered,
much current interest is in the development of purely
algebraic approaches that may be applied in many situations,
without extensive problem-specific tuning.

In this talk,
we present a new algebraic approach to finding the
fine/coarse partitions needed in multilevel approaches. The
algorithm is motivated by theoretical analysis of the
performance of algebraic multigrid (AMG) and algebraic
recursive multilevel solvers (ARMS). From the AMG point of
view, the coarsening is consistent with the ideas of
compatible relaxation, while it may also be motivated by the
algebraic criteria central to ARMS. While no guarantee on
the rate of coarsening is given, the splitting is shown to
always yield an effective preconditioner in the two-level
sense. Numerical performance of two-level and multilevel
variants of this approach is demonstrated.



	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\end{document}

