\documentclass{report}
\usepackage{amsmath,amssymb}

\def\mathbi#1{\textbf{\em #1}}
\def\BA{\bf{A}}
\def\BB{\bf{B}}
\def\bn{\bf{n}}
\def\CP{\mathcal{P}}
\def\CV{\mathcal{V}}
\def\CI{\mathcal{I}}

\newcommand{\gradt}{\nabla\cdot}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\Cplex}{\mathbb{C}}

\newcommand{\Vecc}[2]{ \left(
	\begin{array}{c}
	#1 \\ #2
	\end{array}
	\right) }

\newcommand{\Matrr}[4]{ \left(
	\begin{array}{cc}
	#1 & #2 \\
	#3 & #4
	\end{array}
	\right) }

\newcommand{\calB}{{\mathcal B}}
\newcommand{\calBB}{{\mathcal B}^{\Box}}
\newcommand{\calK}{\mathcal{K}}
\newcommand{\bfA}{{\mathbf A}}
\newcommand{\bfb}{{\mathbf b}}
\newcommand{\bfr}{{\mathbf r}}
\newcommand{\bfx}{{\mathbf x}}
\newcommand{\bfxex}{{\mathbf x}_{\mbox{\scriptsize $\star$}}}
\newcommand{\Dim}{\mathop{\mathrm{dim\ }}}

\begin{document}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{center}
{\large
{\bf Least squares preconditioners for stabilized mixed \\
	approximation of the Navier-Stokes equations}}

	David Silvester \\
	School of Mathematics, University of Manchester \\
	Manchester M60 1QD, UK \\
	{\tt d.silvester@manchester.ac.uk} \\
	Howard Elman, Victoria Howle, John Shadid, Ray Tuminaro
\end{center}
We consider the Navier--Stokes equations
$$
\qquad
\qquad
\qquad
\begin{array}{rcl} -\nu \nabla^2 {\bf u} + ({\bf
u} \cdot {\rm grad})\, {\bf u} + {\rm grad}\, p &= &{\bf f}
\\ -{\rm div}\, {\bf u} &= &0 \end{array}
\qquad
\qquad
\qquad
(1)
$$
on
$\Omega \subset \mathbb{R}^d$, $d=2$ or $3$.
Here, ${\bf u}$
is the $d$-dimensional velocity field, which is assumed to
satisfy suitable boundary conditions on $\partial \Omega$,
$p$ is the pressure, and $\nu$ is the kinematic viscosity,
which is inversely proportional to the Reynolds number.

Linearization and discretization of (1) by finite
elements, finite differences or finite volumes leads to a
sequence of linear systems of equations of the form
$$
\qquad
\qquad
\qquad
\left[
\begin{array}{lc} F & \ B^T \\ B & -\frac{1}{\nu}C
\end{array} \right] \begin{bmatrix} {\bf u}\\ p
\end{bmatrix} = \begin{bmatrix} {\bf f}\\ g \end{bmatrix}.
\qquad
\qquad
\qquad
(2)
$$
These systems, which are the focus of this
talk, must be solved at each step of a nonlinear (Picard or
Newton) iteration. Here, $B$ and $B^T$ are matrices
corresponding to discrete divergence and gradient operators,
respectively and $F$ operates on the discrete velocity
space. For {\em div-stable} discretizations, $C=0$. For
mixed approximation methods that do not uniformly satisfy a
discrete inf-sup condition, the matrix $C$ is a nonzero {\em
stabilization operator}. Examples of finite element methods
that require stabilization are the mixed approximations
using linear or bilinear velocities (trilinear in
three dimensions) coupled with constant pressures, as well
as any discretization in which equal order discrete
velocities and pressures are specified using a common set of
nodes.

The focus of this talk is the Least Squares
Commutator (LSC) preconditioner developed by Elman, Howle,
Shadid, Shuttleworth and Tuminaro, and unveiled at the
Copper Mountain Conference in 2004. This preconditioning
methodology is one of several choices that are effective for
Navier-Stokes equations, and it has the advantage of being
defined from strictly algebraic considerations. The
resulting preconditioning methodology is competitive with
the pressure convection-diffusion preconditioner of Kay,
Loghin and Wathen, and in some cases its performance is
superior. However, the LSC approach has so far only been
shown to be applicable to the case where $C=0$ in
(2). In this talk we show
that the least squares commutator preconditioner can be
extended to cover the case of mixed approximation that
require stabilization. This closes a gap in the derivation
of these ideas, and a version of the method can be also
formulated from algebraic considerations, which enables the
fully automated algebraic construction of effective
preconditioners for the Navier-Stokes equations by
essentially using only properties of the matrices in
(2).

Our focus in this
work is on steady flow problems although the ideas discussed
generalize in a straightforward manner to unsteady flow.


	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\end{document}

