\documentclass{report}
\usepackage{amsmath,amssymb}
\setlength{\parindent}{0mm}
\setlength{\parskip}{1em}
\begin{document}
\begin{center}
\rule{6in}{1pt} \
{\large Katherine Evans \\
{\bf Fully implicit methods for Earth system models}}

Oak Ridge National Laboratory \\ PO Box 2008 \\ MS6016 \\ 1 Bethel Valley Rd \\ Oak Ridge \\ TN 37831
\\
{\tt evanskj@ornl.gov}\\
Damian Rouson\\
Andrew Salinger\\
	James B. White III, Mark Taylor, Roger Pawlowski, Roscoe Bartlett, and John Drake\end{center}

Major algorithmic challenges exist for the solution of Earth system problems,
for example, coupled nonlinear physics, multiple disparate time scales, and
scalability requirements. To address scaling, temporal parallelism is explored
through implementation of the parareal method, which utilizes coarse and fine
time grid operators. The fully implicit (FI) solution method is applied to the
shallow water equations on a sphere as a coarse grid operator, and early results
are presented. The FI solution framework allows relatively large time steps to
be taken stably, within which finer time steps that resolve all the relevant
physics can occur in parallel. The equations are solved within the High-order
Method Modeling Environment (HOMME) model, which uses a cubed sphere grid and
spectral element spatial discretization. The FI solution framework also provides
a coherent nonlinear solution to all dependent variables. Enhanced accuracy with
a second order FI method in time and space is illustrated with a simple fluid
example. The solver implementation is occurring through the development of a
Fortran interface package for the Trilinos project.


\end{document}
