\documentclass{report}
\usepackage{amsmath,amssymb}

\def\mathbi#1{\textbf{\em #1}}
\def\BA{\bf{A}}
\def\BB{\bf{B}}
\def\bn{\bf{n}}
\def\CP{\mathcal{P}}
\def\CV{\mathcal{V}}
\def\CI{\mathcal{I}}

\newcommand{\gradt}{\nabla\cdot}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\Cplex}{\mathbb{C}}

\newcommand{\Vecc}[2]{ \left(
	\begin{array}{c}
	#1 \\ #2
	\end{array}
	\right) }

\newcommand{\Matrr}[4]{ \left(
	\begin{array}{cc}
	#1 & #2 \\
	#3 & #4
	\end{array}
	\right) }

\newcommand{\calB}{{\mathcal B}}
\newcommand{\calBB}{{\mathcal B}^{\Box}}
\newcommand{\calK}{\mathcal{K}}
\newcommand{\bfA}{{\mathbf A}}
\newcommand{\bfb}{{\mathbf b}}
\newcommand{\bfr}{{\mathbf r}}
\newcommand{\bfx}{{\mathbf x}}
\newcommand{\bfxex}{{\mathbf x}_{\mbox{\scriptsize $\star$}}}
\newcommand{\Dim}{\mathop{\mathrm{dim\ }}}

\begin{document}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{center}
{\large
{\bf Novel preconditioning techniques in matrix computations}}

	Victor Pan \\
	Math. and Computer Science Dept. \\
	Lehman College, CUNY, Bronx NY 10468 \\
	{\tt vpan@lehman.cuny.edu}
\end{center}
Multiplicative preconditioners are popular for solving
linear systems of equations. We introduce additive
preconditioners, which are more robust and more readily
available. By combining them with aggregation processes, we
obtain preconditioners that are nearly as effective as
SVD-based multiplicative preconditioners. Besides
simplifying and stabilizing the generation of such
preconditioners, this technique also preserves the structure
of the input matrices, and we also employ it for numerical
computation of the sign of the matrix determinant. Our
alternative techniques employ additive preconditioners for
computing null vectors and null space bases for a singular
matrix, and we extend this approach to yield alternative
linear solvers and to compute the tails of the SVDs of ill
conditioned matrices and matrix eigenvectors. We support our
approach with extensive analysis and numerical experiments.



	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\end{document}

