\documentclass{report}
\usepackage{amsmath,amssymb}

\def\mathbi#1{\textbf{\em #1}}
\def\BA{\bf{A}}
\def\BB{\bf{B}}
\def\bn{\bf{n}}
\def\CP{\mathcal{P}}
\def\CV{\mathcal{V}}
\def\CI{\mathcal{I}}

\newcommand{\gradt}{\nabla\cdot}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\Cplex}{\mathbb{C}}

\newcommand{\Vecc}[2]{ \left(
	\begin{array}{c}
	#1 \\ #2
	\end{array}
	\right) }

\newcommand{\Matrr}[4]{ \left(
	\begin{array}{cc}
	#1 & #2 \\
	#3 & #4
	\end{array}
	\right) }

\newcommand{\calB}{{\mathcal B}}
\newcommand{\calBB}{{\mathcal B}^{\Box}}
\newcommand{\calK}{\mathcal{K}}
\newcommand{\bfA}{{\mathbf A}}
\newcommand{\bfb}{{\mathbf b}}
\newcommand{\bfr}{{\mathbf r}}
\newcommand{\bfx}{{\mathbf x}}
\newcommand{\bfxex}{{\mathbf x}_{\mbox{\scriptsize $\star$}}}
\newcommand{\Dim}{\mathop{\mathrm{dim\ }}}

\begin{document}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{center}
{\large
{\bf A regularized Gauss-Newton method for nonlinear imaging problems \\
	in diffuse optical tomography}}

	Misha Kilmer \\
	Mathematics Dept., Tufts Univ. \\
	503 Boston Ave.,  Medford MA 02155 \\
	{\tt misha.kilmer@tufts.edu} \\
	Eric de Sturler
\end{center}
The goal in diffuse optical tomography for medical imaging
is the joint reconstruction of parameterized images of
absorption and scattering of light in the body. The
reconstruction requires the approximate solution of a
nonlinear least squares problem for the image parameters.
While traditional approaches such as damped Gauss-Newton
(GN) and Levenberg-Marquardt (LM) have been shown to be
effective at solving the imaging problem in its various
forms, a considerable number of additional function and
Jacobian evaluations are involved in determining the correct
step length and/or damping parameter. In 3D imaging
problems, depending on the particular image model, the cost
of one function evaluation is, at a minimum, the cost of a
dense matrix-vector product and in the worst, but more
realistic, case requires the solution of several large-scale
PDEs. A Jacobian evaluation is even slightly more
expensive. Therefore, it is crucial to keep the number of
function and Jacobian evaluations to a minimum.

The
ill-conditioning of the Jacobian, together with the presence
of noise in the data, motivates us to devise a regularized,
trust-region-based Gauss-Newton approach for determining
search directions. Although LM can be thought of as a
regularized analogue to determining the GN direction, LM has
the property of damping possibly important contributions to
the search direction in spectral components corresponding to
small singular values. On the other hand, the Gauss-Newton
direction is too influenced by components due to small
singular values early on, causing the line search to work
hard to refine the step length. We propose a method that
systematically evaluates the potential contribution of each
of the spectral components corresponding to the GN-direction
and constructs the new direction relative to this
contribution within the confines of a trust-region. Examples
show the success of our method in minimizing function
evaluations with respect to other well-known methods.



	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\end{document}

