\documentclass{report}
\usepackage{amsmath,amssymb}

\def\mathbi#1{\textbf{\em #1}}
\def\BA{\bf{A}}
\def\BB{\bf{B}}
\def\bn{\bf{n}}
\def\CP{\mathcal{P}}
\def\CV{\mathcal{V}}
\def\CI{\mathcal{I}}

\newcommand{\gradt}{\nabla\cdot}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\Cplex}{\mathbb{C}}

\newcommand{\Vecc}[2]{ \left(
	\begin{array}{c}
	#1 \\ #2
	\end{array}
	\right) }

\newcommand{\Matrr}[4]{ \left(
	\begin{array}{cc}
	#1 & #2 \\
	#3 & #4
	\end{array}
	\right) }

\newcommand{\calB}{{\mathcal B}}
\newcommand{\calBB}{{\mathcal B}^{\Box}}
\newcommand{\calK}{\mathcal{K}}
\newcommand{\bfA}{{\mathbf A}}
\newcommand{\bfb}{{\mathbf b}}
\newcommand{\bfr}{{\mathbf r}}
\newcommand{\bfx}{{\mathbf x}}
\newcommand{\bfxex}{{\mathbf x}_{\mbox{\scriptsize $\star$}}}
\newcommand{\Dim}{\mathop{\mathrm{dim\ }}}

\begin{document}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{center}
{\large
{\bf On large scale diagonalization techniques for the \\
	Anderson model of localization}}

	Olaf Schenk \\
	Dept.~of Computer Science, University of Basel \\
	Klingelbergstrasse 50, CH-4056, Basel, Switzerland \\
	{\tt olaf.schenk@unibas.ch} \\
	Matthias Bollhoefer, Rudolf Roemer
\end{center}
One of the hardest challenges in modern eigenvalue
computation is the numerical solution of large-scale
eigenvalue problems, in particular those arising from
quantum physics such as the Anderson model of localization
[4]. Typically, these problems require the computation of
some interior eigenvalues and eigenvectors for systems which
have up to several million unknowns due to their high
spatial dimensions. Furthermore, their underlying structure
involves random perturbations of matrix elements which
invalidates simple preconditioning approaches.

We propose
an efficient preconditioning algorithm for this Anderson
model of localization [6]. The model requires the
computation of a few interior eigenvalues and their
associated eigenvectors for large scale, sparse, real and
symmetric indefinite matrices. Our preconditioning approach
for the associated shift-and-invert systems is based on
maximum weighted matchings [3,5] and algebraic multilevel,
inverse-based incomplete $LDL^T$ factorizations [1,2]. Our
numerical examples indicate that recent algebraic multilevel
preconditioning solvers can accelerative the computation of
the underlying large-scale eigenvalue problem by several
orders of magnitude compared with previous approaches [4,6].

[1]
M.~Bollh\"ofer and Y.~Saad,
{\em Multilevel preconditioners
constructed from inverse-based {ILU}s},
SIAM J.~Sci. Comp.,
to appear (2006).

[2]
M.~Bollh\"ofer and O.~Schenk,
{\em {I}{L}{U}{P}{A}{C}{K} volume 2.0 ---
preconditioning software package for
symmetrically structured problems} (May 2005).

[3]
I.~S.~Duff and S.~Pralet,
{\em Strategies for scaling and pivoting for
sparse symmetric indefinite problems},
Technical Report TR/PA/04/59,
CERFACS (2004).

[4]
U.~Elsner, V.~Mehrmann,
F.~Milde, R.~A.~{R\"{o}mer}, M.~Schreiber,
{\em The {Anderson} model of localization: a challenge
for modern eigenvalue methods},
SIAM J.~Sci. Comp.
{\bf 20} (1999) 2089--2102.

[5]
M.~Hagemann, O.~Schenk,
{\em Weighted matchings for the preconditioning
of symmetric indefinite linear systems},
SIAM J.~Sci. Comp.,
to appear (2006).

[6]
O.~Schenk, M.~Bollh\"ofer, R.~R\"omer,
{\em On Large Scale Diagonalization Techniques
For The {Anderson} Model Of Localization},
SIAM J.~Sci. Comp., to appear (2006).



	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\end{document}

