\documentclass{report}
\usepackage{amsmath,amssymb}

\def\mathbi#1{\textbf{\em #1}}
\def\BA{\bf{A}}
\def\BB{\bf{B}}
\def\bn{\bf{n}}
\def\CP{\mathcal{P}}
\def\CV{\mathcal{V}}
\def\CI{\mathcal{I}}

\newcommand{\gradt}{\nabla\cdot}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\Cplex}{\mathbb{C}}

\newcommand{\Vecc}[2]{ \left(
	\begin{array}{c}
	#1 \\ #2
	\end{array}
	\right) }

\newcommand{\Matrr}[4]{ \left(
	\begin{array}{cc}
	#1 & #2 \\
	#3 & #4
	\end{array}
	\right) }

\newcommand{\calB}{{\mathcal B}}
\newcommand{\calBB}{{\mathcal B}^{\Box}}
\newcommand{\calK}{\mathcal{K}}
\newcommand{\bfA}{{\mathbf A}}
\newcommand{\bfb}{{\mathbf b}}
\newcommand{\bfr}{{\mathbf r}}
\newcommand{\bfx}{{\mathbf x}}
\newcommand{\bfxex}{{\mathbf x}_{\mbox{\scriptsize $\star$}}}
\newcommand{\Dim}{\mathop{\mathrm{dim\ }}}

\begin{document}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{center}
{\large
{\bf Block locally optimal preconditioned eigenvalue xolvers (BLOPEX)}}

	Andrew Knyazev \\
	Dept.~of Mathematics, University of Colorado at Denver \\
	P.O.~Box 173364, Campus Box 170, Denver CO 80217-3364 \\
	{\tt andrew.knyazev@cudenver.edu} \\
	Ilya Lashuk, Merico Argentati, Evgueni Ovtchinnikov
\end{center}
Block Locally Optimal Preconditioned Eigenvalue Xolvers
(BLOPEX) is a package, written in C, that at present
includes only one eigenxolver, Locally Optimal Block
Preconditioned Conjugate Gradient Method (LOBPCG). BLOPEX
supports parallel computations through an abstract layer.
BLOPEX is incorporated in the HYPRE package from LLNL and is
available as an external block to the PETSc package from ANL
as well as a stand-alone serial library. HYPRE and PETSc
packages provide high quality multigrid and domain
decomposition preconditioning on parallel clusters with
distributed or shared memory.

 The LOBPCG method,
suggested and developed by Andrew Knyazev [1] in the past
decade, recently attracts an increasing attention as a
potential alternative to the shift-and-invert Lanczos and
preconditioned Davidson methods due to its simplicity,
robustness and fast convergence. Several MATLAB, C, C++ and
FORTRAN implementations of the LOBPCG are developed by
different groups, e. g., for such applications areas as
structured mechanics and electronic structure calculations.
Main LOBPCG features: a matrix-free iterative method for
computing several extreme eigenpairs of symmetric positive
generalized eigenproblems; a user-defined preconditioner;
robustness with respect to random initial approximations,
variable preconditioners, and ill-conditioning of the
stiffness matrix; apparently optimal convergence speed.
Numerical comparisons suggest that LOBPCG may be a genuine
block analog for eigenproblems of the standard
preconditioned conjugate gradient method for symmetric
linear systems.

We present initial scalability results
using BLOPEX with HYPRE and PETSc on one BlueGene/L box
solving eigenvalue problems of record sizes.

[1] A.~V.~Knyazev, {\em Toward the Optimal Preconditioned
Eigensolver: Locally Optimal Block Preconditioned Conjugate
Gradient Method}, SIAM J.~Sci. Comp. {\bf 23}(2) (2001)
517--541.



	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\end{document}

