\documentclass{report}
\usepackage{amsmath,amssymb}

\def\mathbi#1{\textbf{\em #1}}
\def\BA{\bf{A}}
\def\BB{\bf{B}}
\def\bn{\bf{n}}
\def\CP{\mathcal{P}}
\def\CV{\mathcal{V}}
\def\CI{\mathcal{I}}

\newcommand{\gradt}{\nabla\cdot}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\Cplex}{\mathbb{C}}

\newcommand{\Vecc}[2]{ \left(
	\begin{array}{c}
	#1 \\ #2
	\end{array}
	\right) }

\newcommand{\Matrr}[4]{ \left(
	\begin{array}{cc}
	#1 & #2 \\
	#3 & #4
	\end{array}
	\right) }

\newcommand{\calB}{{\mathcal B}}
\newcommand{\calBB}{{\mathcal B}^{\Box}}
\newcommand{\calK}{\mathcal{K}}
\newcommand{\bfA}{{\mathbf A}}
\newcommand{\bfb}{{\mathbf b}}
\newcommand{\bfr}{{\mathbf r}}
\newcommand{\bfx}{{\mathbf x}}
\newcommand{\bfxex}{{\mathbf x}_{\mbox{\scriptsize $\star$}}}
\newcommand{\Dim}{\mathop{\mathrm{dim\ }}}

\begin{document}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{center}
{\large
{\bf Implementation and performance of a two-grid method for \\
	nonlinear reaction-diffusion equations}}

	Jason S.~Howell \\
	Dept.~of Mathematical Sciences \\
	O-106 Martin Hall, Box 340975, Clemson SC 29634-0975 \\
	{\tt jshowel@clemson.edu} \\
	Carol S.~Woodward, Todd S.~Coffey
\end{center}
Two-grid methods have been developed by Xu (SIAM J.~Num.
Anal., 1996) for application to linear and nonlinear PDEs. Of
particular interest are methods that can be applied to large
nonlinear problems that arise in the simulation of physical
processes, such as a method due to Dawson, Wheeler, and
Woodward (SIAM J.~Num. Anal., 1998). This scheme solves the
original nonlinear problem on a mesh coarser than originally
specified to capture the nonlinear behavior of the solution,
then utilizes a linearized version of the problem to correct
the coarse approximation on the original problem mesh.

The
two-grid method potentially reduces the overall
computational cost by requiring the solution of a smaller
nonlinear system and a large linear system in place of the
original large nonlinear problem. In this talk we
investigate the application of this method to nonlinear
reaction-diffusion equations. In particular, we discuss
issues that arise in the implementation of the algorithm,
and perform numerical experiments on problems designed to
gauge the performance of the two-grid method relative to a
standard Newton iterative nonlinear solver.



	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\end{document}

