\documentclass[11pt]{article}
\usepackage{amsfonts}
\setlength{\parskip}{1.2ex}      % space between paragraphs
\setlength{\parindent}{0em}      % amount of indention
\setlength{\textwidth}{165mm}    % default = 6.5in
\setlength{\oddsidemargin}{0mm}  % default = 0mm
\setlength{\textheight}{225mm}   % default = 9in
\setlength{\topmargin}{-1mm}     % default = 0mm
\date{ ~ \hspace{-4mm}}


\title{Calculation of Invariant Subspaces of Large and Sparse Matrices.  }

\author{Jan Brandts \\ {\tt brandts@math.uu.nl} \\ Mathematics Institute, Utrecht University, Budapestlaan 6, 3584 CD Utrecht,  Netherlands.}

\begin{document}
\maketitle
\thispagestyle{empty}





 



Eigendata calculations for large and sparse matrices are mostly done 
by Ritz-Galerkin projection of the large problem onto a small subspace. 
This happens for example in the Power Method (on a one-dimensional 
subspace), in Subspace Iteration and Block Rayleigh Quotient Iteration (on 
a subspace of a fixed dimension p), the Lanczos and the Arnoldi methods, 
and also in the Jacobi-Davidson method (on subspaces of variable dimensions).



In methods in which the subspace is changed or expanded, the question of 
how to do that is a very important one. The ideal change is given by the 
solution of a generalized algebraic Riccati equation, which would yield
an invariant subspace immediately. This Riccati equation, however, is as
hard to solve as the original problem.



In this presentation we will show the consequences of solving the Riccati
equation by iterative methods, and using the iterates as new subspace 
vectors, or as vectors with which to expand a given subspace. This will 
lead to an improvement of Inexact Block Rayleigh Quotient Iteration, and 
also to new algorithms for invariant subspaces that are block versions 
(or variations of those) of the Jacobi-Davidson Method.



We stress that the algorithms for the Riccati equation can be of interest in
other areas of research as well, such as control theory, and in solving the
Ricatti differential equations.



Keywords: Riccati equation, Ritz-Galerkin method, Sylvester equation, stability
of invariant subspaces, Block Jacobi-Davidson method.




\end{document}
