\documentclass[11pt]{article}
\usepackage{amsfonts}
\setlength{\parskip}{1.2ex}      % space between paragraphs
\setlength{\parindent}{0em}      % amount of indention
\setlength{\textwidth}{165mm}    % default = 6.5in
\setlength{\oddsidemargin}{0mm}  % default = 0mm
\setlength{\textheight}{225mm}   % default = 9in
\setlength{\topmargin}{-1mm}     % default = 0mm
\date{ ~ \hspace{-4mm}}


\title{Applications of Brandt's General Algebraic Coarsening  }

\author{David Moulton ({\tt moulton@lanl.gov}) \\ Joel Dendy \\ \\ Mathematical Modeling and Analysis Group  \\  Theoretical Division, Los Alamos National Laboratory  \\  Los Alamos, NM 87545  \\  LA-UR-00-577}

\begin{document}
\maketitle
\thispagestyle{empty}





 



    The development of robust structured multigrid methods, such as
Dendy's black box multigrid, is based on operator-induced
(matrix-dependent) interpolation with variational coarsening.
Similarly, for more general problems, algebraic multigrid employs
variational coarsening.  Although some progress has been made in this
area (e.g., AMGe), controlling the accuracy and growth of the
coarse-grid operators is difficult in this setting.  In contrast
Brandt's general algebraic coarsening is based on a constrained
least-squares formulation.  Similar to algebraic methods it uses
coarse-grid selection as its primary tool, but more importantly, the
least-squares approach offers a flexible formulation for limiting the
growth of coarse-grid operators.




   In this work we consider the application of Brandt's algebraic
coarsening procedure to two applications.  First we consider
homogenization or upscaling of the permeability in porous media, where
we investigate the convergence of Brandt's method as well as compare with
our earlier work on Multigrid Homogenization.  Second we will apply
the method to mixed-discretizations of the diffusion operator,
examining the resulting coarse-grid operators, and the efficiency of
the resulting multilevel method.




This research is supported by the Department of Energy,
  under contract W-7405-ENG-36





\end{document}
