\documentclass{report}
\usepackage{amsmath,amssymb}

\def\mathbi#1{\textbf{\em #1}}
\def\BA{\bf{A}}
\def\BB{\bf{B}}
\def\bn{\bf{n}}
\def\CP{\mathcal{P}}
\def\CV{\mathcal{V}}
\def\CI{\mathcal{I}}

\newcommand{\gradt}{\nabla\cdot}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\Cplex}{\mathbb{C}}

\newcommand{\Vecc}[2]{ \left(
	\begin{array}{c}
	#1 \\ #2
	\end{array}
	\right) }

\newcommand{\Matrr}[4]{ \left(
	\begin{array}{cc}
	#1 & #2 \\
	#3 & #4
	\end{array}
	\right) }

\newcommand{\calB}{{\mathcal B}}
\newcommand{\calBB}{{\mathcal B}^{\Box}}
\newcommand{\calK}{\mathcal{K}}
\newcommand{\bfA}{{\mathbf A}}
\newcommand{\bfb}{{\mathbf b}}
\newcommand{\bfr}{{\mathbf r}}
\newcommand{\bfx}{{\mathbf x}}
\newcommand{\bfxex}{{\mathbf x}_{\mbox{\scriptsize $\star$}}}
\newcommand{\Dim}{\mathop{\mathrm{dim\ }}}

\begin{document}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{center}
{\large
{\bf Development of a 2-D model to simulate convection and \\
	phase transition efficiently}}

	Katherine J.~Evans \\
	MS B216, PO Box 1663 \\
	Los Alamos National Laboratory, Los Alamos NM 87545 \\
	{\tt kevans@lanl.gov} \\
	D.~A.~Knoll, Michael Pernice
\end{center}
We present a two-dimensional convection phase change model
using the incompressible Navier-Stokes equation set and
enthalpy as the energy conservation variable. Significant
algorithmic challenges are posed for problems of phase
change interfaces within convective flow regimes. The
equation set is solved with the Jacobian-Free Newton-Krylov
(JFNK) nonlinear inexact Newton's method. SIMPLE, a
pressure-correction algorithm, is used as a physics-based
preconditioner. This algorithm is compared to solutions
using SIMPLE as the main solver.

Algorithm performance is
assessed for a benchmark problem, phase change convection
within a square cavity of a solid pure material cooled below
the melting temperature. A time step convergence analysis
demonstrates that the JFNK model with second order
discretization is second order accurate in time. A Gallium
melting simulation is also performed and evaluated; in this
configuration multiple roll cells develop in the melted
region at early times when the aspect ratio is high. The
JFNK-SIMPLE method is shown to be more efficient per time
step and more robust at larger time steps when compared to
SIMPLE as the main solution algorithm. Overall CPU savings
of more than an order of magnitude are realized.

As a
further analysis of JFNK-SIMPLE, multigrid is wrapped around
SIMPLE, so SIMPLE acts as the smoother within a multigrid
preconditioner to JFNK. For phase change conduction,
additional gains in efficiency can be accomplished. When
SIMPLE is incorporated as a preconditioner and smoother
within JFNK, the ability to model more complex and realistic
phase change convection problems with increased robustness
and efficiency is achieved.



	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\end{document}

