\documentclass{report}
\usepackage{amsmath,amssymb}
\setlength{\parindent}{0mm}
\setlength{\parskip}{1em}
\begin{document}
\begin{center}
\rule{6in}{1pt} \
{\large J. Mas \\
{\bf A Balanced Incomplete Factorization preconditioner}}

Institut de Matem\`atica Multidisciplinar \\ Universitat Polit\`ecnica de Val\`encia \\ Cam\'{\i} de Vera \\ 14 \\ 46022 Val\`encia \\ Spain
\\
{\tt jmasm@imm.upv.es}\\
R. Bru\\
J. Mar\'{\i}n\\
	T\accent23uma, M.\end{center}

In~\cite{bcmm:03} a factorized approximate inverse preconditioner
(AISM) based on the Sherman-Morrison formula is presented. In this work
the structure of the factors is analyzed and it is proved that the $L$ and
$D$ factors of the $LDU$ factorization of a matrix are easily retrieved from
one of them.
Moreover columns of $L$ and $L^{-1}$ are computed in the same step which makes
the application of dropping strategies proposed in~\cite{bs:02} straightforward
in the case of a symmetric positive definite coefficient matrix.

Numerical experiments on a set of symmetric positive definite matrices are done
showing that this strategy, that we called Balanced Incomplete Factorization
(BIF for short), is very robust for preconditioning very ill--conditioned
problems using iterative methods.
In addition BIF exhibits shorter setup times than RIF \cite{betu:03},
a method
of a similar and very high level of robustness.


\begin{thebibliography}{99}
\bibitem{betu:03}
M.\ Benzi and M.\ T\accent23uma.
{\em A robust incomplete factorization preconditioner for positive definite
matrices}, Numer.\ Linear Algebra Appl., 10(5-6), (2003) 385--400, .
\bibitem{bs:02}
M.\ Bollh\"ofer and Y.\ Saad.
{\em On the relations between ILUs and factored approximate inverses}.
SIAM J.\ Matrix Anal.\ Appl., 24(1) (2002) 219--237.
\bibitem{bcmm:03}
R.\ Bru, J.\ Cerd\'an, J.\ Mar\'{\i}n and J.\ Mas, {\em Preconditioning
sparse nonsymmetric linear systems
with the Sherman-Morrison formula}, SIAM J.\ Sci.\ Comput., 25(2) (2003),
701--715.

\end{thebibliography}


\end{document}
