\documentclass{report}
\usepackage{amsmath,amssymb}

\def\mathbi#1{\textbf{\em #1}}
\def\BA{\bf{A}}
\def\BB{\bf{B}}
\def\bn{\bf{n}}
\def\CP{\mathcal{P}}
\def\CV{\mathcal{V}}
\def\CI{\mathcal{I}}

\newcommand{\gradt}{\nabla\cdot}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\Cplex}{\mathbb{C}}

\newcommand{\Vecc}[2]{ \left(
	\begin{array}{c}
	#1 \\ #2
	\end{array}
	\right) }

\newcommand{\Matrr}[4]{ \left(
	\begin{array}{cc}
	#1 & #2 \\
	#3 & #4
	\end{array}
	\right) }

\newcommand{\calB}{{\mathcal B}}
\newcommand{\calBB}{{\mathcal B}^{\Box}}
\newcommand{\calK}{\mathcal{K}}
\newcommand{\bfA}{{\mathbf A}}
\newcommand{\bfb}{{\mathbf b}}
\newcommand{\bfr}{{\mathbf r}}
\newcommand{\bfx}{{\mathbf x}}
\newcommand{\bfxex}{{\mathbf x}_{\mbox{\scriptsize $\star$}}}
\newcommand{\Dim}{\mathop{\mathrm{dim\ }}}

\begin{document}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{center}
{\large
{\bf A novel deflation method to solve the 3-D discontinuous \\
	and singular Poisson equation}}

	Jok M.~Tang \\
	Dept.~of Applied Mathematical Analysis \\
	Delft University of Technology \\
	Mekelweg 4, 2628 CD Delft, The Netherlands \\ 
	{\tt j.m.tang@ewi.tudelft.nl}
\end{center}
Recently, simulating bubbly flows is a very popular topic in
CFD. These bubbly flows are governed by the Navier-Stokes
equations. In many popular operator splitting formulations
for these equations, solving the linear system coming from
the discontinuous Poisson equation takes the most
computational time, despite of its elliptic origins. ICCG is
widely used for this purpose, but for complex bubbly flows
this method shows slow convergence.

Moreover, new insights are given into the properties of
invertible and singular deflated and preconditioned linear
systems, where the coefficient matrices are symmetric and
positive (semi-) definite. These linear systems can be
derived from a discretization of the Poisson equation with
Neumann boundary conditions. Sometimes these linear systems
are forced to be invertible leading to a worse (effective)
condition number. If ICCG is used to solve this problem, the
convergence is significantly slower than for the case of the
original singular problem.

We show that applying the deflation technique, which leads
to the DICCG method, remedies the worse condition number and
the worse convergence of ICCG. Moreover, some useful
equalities are derived from the deflated variants of the
singular and invertible matrices, which are also generalized
to preconditioned methods. It appears that solving the
invertible and singular linear systems with DICCG leads to
exactly the same convergence results.

This new method DICCG incorporates the eigenmodes
corresponding to the components which caused the slow
convergence of ICCG. Coarse linear systems have to be solved
within DICCG. We discuss some methods to do this efficiently
which results in two approaches DICCG1 and DICCG2.

Thereafter we show with numerical experiments that both DICCG
approaches are very efficient and they emphasize also the
theoretical results. Compared to ICCG, DICCG decreases
significantly the number of iterations and the computational
time as well which are required for solving Poisson equation
in applications of 2-D and 3-D bubbly flows.



	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\end{document}

