\documentclass{report}
\usepackage{amsmath,amssymb}

\def\mathbi#1{\textbf{\em #1}}
\def\BA{\bf{A}}
\def\BB{\bf{B}}
\def\bn{\bf{n}}
\def\CP{\mathcal{P}}
\def\CV{\mathcal{V}}
\def\CI{\mathcal{I}}

\newcommand{\gradt}{\nabla\cdot}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\Cplex}{\mathbb{C}}

\newcommand{\Vecc}[2]{ \left(
	\begin{array}{c}
	#1 \\ #2
	\end{array}
	\right) }

\newcommand{\Matrr}[4]{ \left(
	\begin{array}{cc}
	#1 & #2 \\
	#3 & #4
	\end{array}
	\right) }

\newcommand{\calB}{{\mathcal B}}
\newcommand{\calBB}{{\mathcal B}^{\Box}}
\newcommand{\calK}{\mathcal{K}}
\newcommand{\bfA}{{\mathbf A}}
\newcommand{\bfb}{{\mathbf b}}
\newcommand{\bfr}{{\mathbf r}}
\newcommand{\bfx}{{\mathbf x}}
\newcommand{\bfxex}{{\mathbf x}_{\mbox{\scriptsize $\star$}}}
\newcommand{\Dim}{\mathop{\mathrm{dim\ }}}

\begin{document}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{center}
{\large
{\bf Algebraic multigrid and convection diffusion problems}}

	Jonathan W Boyle \\
	School of Mathematics, The University of Manchester \\
	Sackville Street, Manchester M60 1QD UK \\
	{\tt j.boyle@manchester.ac.uk} \\
	David J Silvester
\end{center}
The talk discusses the use of algebraic multigrid (AMG) when
solving discrete convection diffusion problems. For
practical applications, e.g. modeling incompressible fluid
flow, the interest is in using AMG as a preconditioning
component within a Krylov subspace solver like GMRES. To be
effective in this role, the AMG method should be 'black
box', i.e. require no user tuning, and be effective in the
sense that if applied as a solver for a convection-diffusion
subproblem then iteration counts are independent of mesh
size and other problem parameters.

When designing AMG for
convection diffusion, it is necessary to consider the
interplay between wind direction and AMG smoothing scheme.
Numerical test data is presented showing performance for AMG
applied to streamline diffusion stabilized finite element
discretizations of model problems. The results demonstrate
the crucial importance of choosing an appropriate smoothing
scheme. The performance of a ``well-chosen'' AMG method in the
context of pressure-convection diffusion preconditioning of
the Navier-Stokes equations will also be discussed.

This is ongoing work.
It is funded by EPSRC grant EP/C000528/1.


	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\end{document}

