\documentclass{report}
\usepackage{amsmath,amssymb}

\def\mathbi#1{\textbf{\em #1}}
\def\BA{\bf{A}}
\def\BB{\bf{B}}
\def\bn{\bf{n}}
\def\CP{\mathcal{P}}
\def\CV{\mathcal{V}}
\def\CI{\mathcal{I}}

\newcommand{\gradt}{\nabla\cdot}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\Cplex}{\mathbb{C}}

\newcommand{\Vecc}[2]{ \left(
	\begin{array}{c}
	#1 \\ #2
	\end{array}
	\right) }

\newcommand{\Matrr}[4]{ \left(
	\begin{array}{cc}
	#1 & #2 \\
	#3 & #4
	\end{array}
	\right) }

\newcommand{\calB}{{\mathcal B}}
\newcommand{\calBB}{{\mathcal B}^{\Box}}
\newcommand{\calK}{\mathcal{K}}
\newcommand{\bfA}{{\mathbf A}}
\newcommand{\bfb}{{\mathbf b}}
\newcommand{\bfr}{{\mathbf r}}
\newcommand{\bfx}{{\mathbf x}}
\newcommand{\bfxex}{{\mathbf x}_{\mbox{\scriptsize $\star$}}}
\newcommand{\Dim}{\mathop{\mathrm{dim\ }}}

\begin{document}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{center}
{\large
{\bf Quasi-Newton preconditioners for the iterative solution \\
	of nonlinear equation in porous media}}

	Rafael Bru \\
	Inst. Matematica Multidisciplinar, Dept. Matematica Aplicada \\
	Univ. Politecnica Cami de Vera s/n 46021, Valencia, Spain \\
	{\tt rbru@mat.upv.es} \\
	Luca Bergamaschi, Andrea Comerlati,
	Angeles Mart\'{\i}nez and Mario Putti
\end{center}
In this work preconditioners for solving the linear systems
of the Newton method in each nonlinear iteration are
studied. The preconditioners are defined by means of a
Broyden-type rank-one update at each nonlinear iteration, as
described in [1]. We report numerical results of the
application of this approach for the solution of the
nonlinear system of algebraic equations arising from the
finite element discretization of two-phase flow model in
porous media. Sequential and parallel results show that the
efficiency of the proposed method. The parallel version of
the algorithm uses the FSAI approximate inverse as initial
preconditioner [2]. The obtained performances show that the
high cost of the FSAI evaluation is amortized by the
efficiency of the subsequent Broyden updates.

[1]
L.~Bergamaschi, R.~Bru, A.~Martinez and M.~Putti,
{\em Quasi-Newton Preconditioners for the Inexact Newton
Method}, to appear in Electr. Trans. on Num. Analysis,
2006.

[2] L.~Yu. Kolotilina and A.~Yu.~Yeremin,
{\em Factorized sparse approximate inverse preconditionings.
I. Theory},
SIAM J.~Matrix Anal. Appl., {\bf 14} (1993) 45--58.


	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\end{document}

