\documentclass{report}
\usepackage{amsmath,amssymb}

\def\mathbi#1{\textbf{\em #1}}
\def\BA{\bf{A}}
\def\BB{\bf{B}}
\def\bn{\bf{n}}
\def\CP{\mathcal{P}}
\def\CV{\mathcal{V}}
\def\CI{\mathcal{I}}

\newcommand{\gradt}{\nabla\cdot}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\Cplex}{\mathbb{C}}

\newcommand{\Vecc}[2]{ \left(
	\begin{array}{c}
	#1 \\ #2
	\end{array}
	\right) }

\newcommand{\Matrr}[4]{ \left(
	\begin{array}{cc}
	#1 & #2 \\
	#3 & #4
	\end{array}
	\right) }

\newcommand{\calB}{{\mathcal B}}
\newcommand{\calBB}{{\mathcal B}^{\Box}}
\newcommand{\calK}{\mathcal{K}}
\newcommand{\bfA}{{\mathbf A}}
\newcommand{\bfb}{{\mathbf b}}
\newcommand{\bfr}{{\mathbf r}}
\newcommand{\bfx}{{\mathbf x}}
\newcommand{\bfxex}{{\mathbf x}_{\mbox{\scriptsize $\star$}}}
\newcommand{\Dim}{\mathop{\mathrm{dim\ }}}

\begin{document}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{center}
{\large
{\bf Space-time solution of large-scale PDE applications}}

	Andrew Salinger \\
	Sandia National Labs, PO Box 5800 \\
	MS-1111, Albuquerque NM 87185 \\
	{\tt agsalin@sandia.gov} \\
	Daniel Dunalvy, Eric Phipps
\end{center}
Software and algorithms are being developed to efficiently
formulate and solve transient PDE problems as ``steady''
problems in a space-time domain. In this way, sophisticated
design and analysis tools for steady problems, such as
continuation methods, can be brought to bear on transient
(and eventually periodic) problems. This new capability is
being developed in the Trilinos solver framework, and is
designed to present a simple interface to application codes.
The software allows for parallelism over both the space and
time domains.

The main hurdle to make this approach
viable is to be able to efficiently solve the very large
Jacobian matrix for very large the space-time system, with
its characteristic structure. We will present results for a
number of preconditioners and solution methods that we have
developed for this linear system. Numerical results for a
PDE reacting flow application will be presented. These
results will shed some light on the underlying question of
whether it can pay to parallelize the time domain.



	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\end{document}

